% Auto-generated from 43-Network-Reorganization.md
% POV: Aisen | Landscape: City | Location: The Academy

By the fourth month of institutional replacement phase, provincial resistance networks had successfully reorganized themselves to operate simultaneously within and against Covenant institutional structures. They had personnel positioned in educational administration, healthcare facilities, local governance offices, resource allocation systems.

They weren't in positions of authority—that would have been too obvious and would have triggered Covenant security scrutiny. But they were in positions where they could influence how institutions actually functioned at operational level, which is where real power often concentrated in institutional systems.

Meanwhile, Aisen was attempting to reconnect with the Academy network, which had survived Covenant Intelligence investigation through going largely dormant during military occupation phase. Vex had returned to the Academy after transitioning through provincial resistance operations. Caris had remained at the Academy, maintaining minimal intelligence operations while appearing to be simply a dedicated academic administrator.

The Academy network was smaller now—several members had been arrested or killed during investigation phase. But it was also more resilient because it had been forced to develop better operational security protocols.

"What's the Academy network's current capacity?" Aisen asked during one of his infrequent visits back to the Academy under cover of educational research coordination.

"Approximately forty percent of peak operational capacity," Vex reported. "We lost Cynthia to detention. We lost three field operatives to security investigations. We lost access to Archive research materials when the Covenant restricted scholar access during military operations."

"Can you rebuild?" Aisen asked.

"Over time, probably," Vex said. "But it's going to take years, not months. The Covenant security apparatus is sophisticated enough that they damaged our operational infrastructure significantly."

"Is institutional subversion strategy still viable from Academy position?" Aisen asked.

"More viable than open resistance ever was," Vex said. "The Academy is still the Covenant's primary educator training system. If resistance network members can position themselves in educational administration and influence curriculum, the institutional impact extends across generations of future institutional officials."
