% Auto-generated from 12-First-Operations.md
% POV: Aisen | Landscape: City | Location: The Market Ward

The east dormitory's rear courtyard was a maintenance space — a paved rectangle approximately eight metres by twelve with a work-shed along the south wall and two lidded water cisterns bracketing the north entrance. No student-facing function. No seating. Aisen had identified it in his fourth day because it appeared in the shadow map Cynthia had given him as a dead zone in the Academy's observation grid, which meant it was either genuinely unmonitored or it was trap geography. He had been using it approximately every three nights since, just to build a behavioral baseline: if anyone was watching the space, the pattern of his visits would look like an insomniac taking the same night walk until the visits were no longer notable.

By his second week, no one appeared to be watching.

Vex was there on the night he judged it safe to make the overture.

She arrived four minutes before him. He would not have known this from her posture — she was standing near the cistern looking at the wall in the way of someone doing nothing specific — but her shoes were slightly damp, which meant she had come through the maintenance path from the north rather than the courtyard entrance and had therefore been here long enough for the morning dew to settle on her shoe leather.

He stood by the shed and said nothing.

She said nothing.

After approximately ninety seconds she said: "You could have identified yourself in the library last week. More efficient."

"I didn't know which library visit you were tracking," Aisen said.

"The one three days ago. When you were reading administrative history. You spent six minutes on a two-page spread that standard readers finish in under two."

"What told you I wasn't just a slow reader?"

"You don't read slowly in class." She finally looked at him. She was approximately his height, with the posture of someone trained to carry weight on the shoulders rather than the lower back — which was a specific martial training indicator, not a natural postural habit — and her hands were at rest in the specific loose position that was taught in close-quarters training for rapid deployment. He had been taught the same hand position and had taken three months to ingrain it. "You had a question you were thinking about while the page was open. So did I, in the same section, four days earlier."

"What was your question?" Aisen asked.

"Why the Covenant's historical administrative archive doesn't contain the standard treaty documents from the forty-third year," she said. "Which is a gap that's visible if you know what should be there."

He knew what should be there. He knew that it wasn't. He had been noting the gap as part of the communication-system observation work Cynthia had assigned him. This was either a genuine shared line of inquiry or it was a very well-prepared probe.

"Who trained you?" he asked — the direct question, which was also the one that would resolve fastest whether she was genuine or prepared.

She was quiet for exactly the amount of time that suggested she was making a real decision rather than consulting a prepared answer. "Someone named Kael. She works at the outer administrative level of the Academy. Not who you'd identify as resistance from the outside." She said it with the quality of someone who had thought about this discrepancy for some time and not yet resolved it. "She approached me two years before I enrolled. She said I had certain qualities that made me potentially useful. She was right about that. I'm less certain about what she's eventually asking me to be useful for."

"What do you mean?"

"I mean she trained me toward disruption. Making systems spend resources maintaining control rather than exercising it. She didn't train me toward the same thing you're doing." She paused. "Whatever you're doing."

Aisen thought about what to say. He was standing in a dead zone at three in the morning with someone who had already demonstrated she could read him, and the decision was: give her a calibrated partial truth or give her nothing and lose the possibility of whatever she was. His chest was doing the unnamed-sensation thing. He noticed it and held still.

"I was trained toward information gathering," he said. "Understanding how the system operates so that eventually someone who knows what they're doing has enough structural knowledge to act on."

"Toward what?"

"I don't know the specific endpoint," he said honestly. "I was told I'd recognize the moment when it existed."

Vex considered this. "So you're working with incomplete information provided by someone who's decided you'll understand the rest from context when context exists."

"Yes," he said. "Is that different from your situation?"

She made the specific quiet sound that was not quite a laugh. "No," she said. "It is not."

A pause. Then Vex glanced at his hands. "Your left thumb. You're still tucking it. That's the three-month version of the hold."

Aisen adjusted without thinking. She nodded once — not approval, acknowledgment. "Better. You were going to figure it out eventually, but eventually is slow."

It was the closest thing to a friendly gesture anyone had offered him since arriving at the Academy, and it came in the form of a tactical correction delivered at three in the morning in a dead zone. He filed that under observations he did not yet have a category for.

They stood in the courtyard for another twenty minutes. He did not ask for her network's details and she did not ask for his, and at some point in the conversation this implicit agreement had been established without either of them explicitly establishing it. What they exchanged was method: how she tracked observation patterns, how he mapped communication flow, what they had each seen in their first two weeks that seemed anomalous against the Academy's stated functions. They were comparing notes in a language that required no translation because it was exactly the same language, applied from parallel angles to the same object.

Before she left, she said: "Same time, three nights from now. If either of us doesn't appear, the other doesn't look for a reason."

"Agreed," he said.

She left north through the maintenance path. He stayed for twelve minutes and then left east.

The next six weeks produced what Cynthia had asked for: a mapping of the Academy's communication architecture from the observable surface inward. Aisen tracked which administrative offices received courier visits at which intervals, which scholars' movements were accompanied by the specific half-second posture adjustment of people carrying sensitive materials, which maintenance staff were on schedules that didn't align with maintenance needs. He wrote nothing down in the field. He retained it and wrote in merchant cipher in his notebook after each session, in the compressed notation that could fit a day's operational observation in three lines.

Vex's observations filled in angles Aisen couldn't cover from his position — she had access to the combat track's administrative flow, which processed through different channels than the academic track's, and she had tracked three staff members whose visible function was scheduling and whose actual movement pattern suggested a secondary role in the information system.

They compared findings in the courtyard. Neither of them reported the other to their respective networks. Neither of them said explicitly that they weren't reporting; this was simply a condition they had mutually established without discussing.

He told Caris nothing about Vex. He was aware this was a deliberate deception of people he was supposed to be working with. He was also aware that introducing Vex to the network before he understood the network's full scope was the faster route to an outcome he had not yet been given enough information to evaluate. He filed the deception under \emph{active unknowns: tolerated ambiguity} in his private taxonomy.

The meeting with Caris occurred through a deliberate spilled drink at the dining hall — a standard brushpast that transferred a small folded paper — which directed Aisen to a supply room in the western dormitory basement. He arrived to find it smelled of cleaning compound and was lit by a single tallow candle that Caris had brought himself.

Caris was approximately forty-five, with the kind of physical economy that came from a long time in operational work — nothing was wasted, each movement serving an exact function. He reviewed the coded notes Aisen handed him with the attention of someone who had learned to read this cipher and was genuinely reading it, not performing a review for visual effect.

"You identified the secondary channel," Caris said. He was not praising it — he was confirming it with the neutrality of someone checking a measurement. "The one that runs through the maintenance staff."

"Yes," Aisen said. "Three staff members. I couldn't determine the content of what they're carrying, only the routing pattern."

"Content isn't necessary for this phase." He set the notes on the cistern surface and looked at them without touching them further. "The next step is the Archive. We need someone inside, not to extract materials, but to document the organizational system. What section headings exist. What security tier each section carries. How materials are retrieved." He paused. "Cynthia has a research approval that grants her access to historical records. She's going to request a student assistant. You're going to volunteer."

Aisen looked at him. "Why me rather than someone already familiar with the Archive's layout?"

"Because no one we have is familiar with it in the capacity that's useful," Caris said. "And because you've built six weeks of documented research interest that gives you a plausible reason to be Cynthia's assistant. An assignment request from someone with no prior research interests would draw scrutiny. You've been in the library regularly; you've asked archival questions in two classes; you've documented your interest in pre-Covenant administrative history. The student profile supports the access."

He had been building the profile, Aisen understood, without knowing it was being built for this purpose. Or — he corrected himself — he had been building it knowing network work required positions, and the specific connection between the observation profile and Archive access had not been made explicit to him until now.

"What are the failure modes?" Aisen asked.

"The Archive has human review of all student access. Someone will notice if you access materials outside Cynthia's research scope. So you stay within scope, you complete the documentation of what you can see from Cynthia's assigned zones, and you retain the structural observations rather than recording them."

"How many sessions will it take?"

"Four to six. Each one is thirty to forty minutes of Archive time. Cynthia will provide cover for your presence on each visit." Caris rolled the notes and held them over the candle until they caught, then dropped them in the cleaning compound bucket where they burned cleanly. "Are you ready for this?"

Aisen considered the question in the way it deserved to be considered, which was as a genuine question rather than a formality. Was he ready. He had surveillance training, observation practice, six weeks of Academy context. He did not have Archive experience or knowledge of what the security felt like from inside. He did not know what he would encounter that Caris's planning had not modeled.

The version of readiness that meant \emph{no residual unknowns} did not exist. The version that meant \emph{enough preparation to operate} — that one he could make an honest assessment of.

"Yes," he said.

Caris nodded once — the confirmation nod, not the approval nod — and the meeting was over. He left first. Aisen counted to two hundred and left second, carrying the smell of cleaning compound and burned paper with him up the stairs into the corridor and the night.

That night, he told Vex what was happening.

They sat in the courtyard, and Aisen said the plan to infiltrate the Archive. Vex listened without interruption, processing the information with the kind of focused attention that suggested she was already calculating implications.

"That's dangerous," she said when he finished.

"Probably," Aisen agreed.

"But it's also strategic," Vex continued. "Information about how the Covenant actually stores and protects their knowledge—that's the kind of thing Kael is trying to build toward. Different approach, same target."

She was quiet for a long moment.

"I'm going to talk to Kael about whether our networks might want to coordinate," she said. "Not merge—I don't think that's realistic. But share information about strategy and timing. Because if two separate networks are trying to destabilize the Covenant independently, we're going to step on each other's operations."

"Is that possible?" Aisen asked. "Network coordination?"

"I don't know," Vex said. "But we should find out. Because if Kael's approach to destabilization and your network's information gathering could be synchronized, that would be much more effective than both of us working in isolation."

It was a risk—revealing the network's existence to another network's operative. But Aisen suspected Caris and Cynthia would see the same strategic value that Vex had identified. And having Vex as an active ally at the Academy, someone positioned to coordinate with another resistance structure, was worth the risk of their knowledge spreading within underground networks.

"Let me know what Kael says," Aisen said.

"And let me know how the Archive operation goes," Vex replied. "Because if you get actual evidence of what the Covenant is protecting in their restricted sections, that's something both networks need to know about."

The courtyard fell silent around them, the Academy sleeping while both of them remained awake, carrying plans and secrets that extended far beyond the institutional walls that supposedly contained them.
