% Auto-generated from Chapter-12-The-Amara-Caravan-Meeting.md
% Source: C:\Users\PCGAME\Desktop\story&universes\chain-weapon-story\finished-manuscript\manuscriptbaseforchapters\Chapter-12-The-Amara-Caravan-Meeting.md

\chapter{The Amara Caravan Meeting}
\renewcommand{\CurrentChapterNum}{12}
\POV{\glyphAisen}{Aisen}

The caravan arrived on the last day of the autumn trade window, when the light had shortened to the point where the afternoon felt like an apology for the morning and the evenings came on with the swiftness of a door being closed. Aisen heard it before he saw it—the particular percussion of a large convoy moving through the valley road: harness brass jingling in compound rhythm, axle-groan from heavy-laden wagons, the bass-note thudding of draft horses whose hooves had been wrapped in leather against the season's mud. Beneath these, the human sounds: voices calling distances, a drover's whistle pitched to cut through wind, the murmured negotiations of outriders sorting right-of-way with local traffic.

He was behind the bar, cleaning glasses. The Crossing's evening crowd had not yet arrived, and the common room existed in that suspended hour between the day's last travelers and the night's first regulars, the room holding itself in the amber pause of tallow light and the smell of the stew that simmered permanently in the kitchen's iron pot—mutton, root vegetables, and the rosemary that grew wild along the riverbank and arrived in the kitchen via Aisen's habit of collecting useful things. His hands worked the cloth around the inside of a glass while his ears parsed the approaching sounds with the unconscious fluency of someone who had grown up at a crossroads and learned to read arriving traffic the way others read weather.

This caravan was larger than the usual convoys. The sound-signature was wrong for a standard merchant train—too many wheels, too many hooves, the spacing between vehicles suggesting professional formation rather than the loose clustering of independent traders sharing a road for safety. Aisen set down the glass, wiped his hands on his apron, and went to the window.

Twelve wagons. He counted them as they appeared around the bend in the valley road, each one canvas-topped and heavy on its springs, drawn by pairs of the sturdy mountain horses that came from the northern breeding grounds. Outriders flanked the column—not militia, not Covenant, but private guards wearing the muted gray-green of professional security, their weapons visible but not brandished. A lead wagon flew a merchant's pennant: three interlocking circles on a field of deep blue. Aisen recognized the sigil from his father's business conversations—the Osei Trading Confederacy, a network that ran goods along the southern routes and had a reputation for carrying more information than inventory.

Harald appeared from the kitchen, drying his hands on a towel. He looked through the window, read the same details Aisen had read, and his expression shifted by increments that a stranger would not have noticed: a slight tightening around the eyes, a fractional lift of the chin, the micro-adjustments of a man recalibrating his evening around a new variable.

"Twelve wagons means they'll need the yard and the overflow stable," Harald said. "Start the second stew pot."

Aisen started the second stew pot. But his attention remained at the window, where the caravan was now pulling into The Crossing's yard with the choreographed efficiency of people who had done this many times before, and among the figures descending from wagons and unhitching horses, one in particular caught his eye and held it.