% Auto-generated from 05-Exam-Sabotage.md
% POV: Aisen | Landscape: Mountains | Location: The Grey Mountains

The examination hall smelled of tallow wax, instrument oil, and the specific sour quality of forty people's collective anxiety concentrated by a low ceiling and insufficient ventilation. Aisen had been standing at the left edge of the gallery for eleven minutes. He was not a candidate — at seventeen, he had no formal standing with the Academy — but Master Wei had issued an observation permit, which meant he had paper authorization to be here and no authorization to do anything except watch.

He was watching.

The resonance benches were arranged in a half-moon facing the gallery's raised seats. Twelve benches, twelve candidates, twelve control rings laid on velvet pads with the precision of objects that had been cleaned and positioned and cleaned again. The reserve crystals in their housings at the left side of each bench were the kind that cost more than Aisen's entire building's combined annual rent, which was relevant because it meant the Academy's procurement records for them would be meticulous, and meticulous records were exactly the kind of record that became useful when something went wrong.

The first candidate tied his throat ribbon — the formal acknowledgment that the trial was beginning — and set the ring against the crystal housing.

Standard procedure: the ring should respond to the crystal's baseline harmonic within four seconds, pulse twice, and then settle into the steady low bob that indicated correct phasing. Patrons in the gallery would note the settling time; the examining masters would note the ring's stability under prolonged hold; the recorder crystal at the front of the bench would capture the harmonic signature for the official result.

The ring pulsed.

Then the bench shivered.

It was a small motion — barely two millimetres of lateral displacement — but in a calibration environment constructed to eliminate all vibration, two millimetres was an event. The ring's feedback fractured, a sound that registered at the edge of Aisen's hearing rather than above it, a half-frequency that raised the small hairs along the back of his right forearm before he understood what he was hearing.

The student — his name was Dorin, Aisen had seen him in the corridor this morning walking with the tightly held posture of someone managing an emotion he had decided not to display — looked up from the bench with an expression that was trying very hard to be professional and was not quite succeeding. His hands had gone still. He was waiting for something to resolve in a way it was not going to resolve.

The crystal's light went out.

The examination hall's collective breath went out with it — a single expiration from forty people, which took approximately two seconds and was followed by the specific silence of a room that didn't know which of the available bad outcomes it was currently experiencing.

Aisen was already moving before the silence fully settled. Not quickly — walking quickly across an examination floor was exactly the behaviour that would get his observation permit revoked — but steadily, toward the front bench nearest the sealed reserve crystal crate at the side wall. He reached it before the Provost's steward did, which meant he had approximately forty-five seconds before official hands arrived and the object was transferred to institutional custody and therefore beyond informal examination.

He crouched by the bench's base.

The smell hit him first: wet earth and something else underneath — not quite sulphur, not quite iron, the particular dry-mineral quality of a crystalline compound that had been recently activated. He had read about it. He had never smelled it in person. His chest did a brief uncomfortable thing.

\emph{Disruption paste.} Used in the third century to interfere with mage-band communication before the practice was classified and all known formulations were added to the prohibited materials registry. The knowledge was in the Archives' restricted section in a volume that was specifically not indexed under the substance's modern name.

Someone had used it on the bench.

He ran his thumb along the inner lip of the crystal housing, using the flat of his fingernail rather than the pad — fingerprints were irrelevant here but nail-contact wouldn't contaminate the residue — and felt the etching before he saw it: a hairline scratch, regular, geometric, not accidental. Three millimetres long, running parallel to the housing's bottom edge. The scratch would have been made to apply the paste, which was typically solid-form and needed a surface disruption to bond.

He had exactly fifteen seconds before the steward arrived.

He looked at the crate.

The crate's seal was intact. That wasn't the information he needed. He needed the inventory log that would be sitting in the steward's ledger on the front table, specifically the late-entry log for any access to the reserve crate in the past seventy-two hours. But the ledger was on the table, the steward was crossing toward him, and the Provost's voice had just said, from the gallery above, something with the word "sealed" in it that had the declarative rhythm of an instruction rather than a question.

The benches were going to be closed.

He had to get to the ledger before they sealed the benches, because sealing the benches meant transferring custody to the Covenant's examination oversight committee, which meant the log became restricted material, which meant anyone looking at it would need institutional authorization, which meant the two or three days that would pass before that authorization was granted would be two or three days in which the narrative about what had happened here would harden around whoever the Provost decided was most convenient.

"I need to see the late-entry log," Aisen said to the steward, before the steward had fully arrived at the bench.

The steward was a man in his forties whose primary characteristic was a studied blankness — the face of someone who had organized their expression to reveal nothing, which Aisen had learned meant either genuine composure or high anxiety, and you determined which by looking at the hands. His hands were holding the bench inventory clipboard with slightly more force than the weight of the clipboard required.

High anxiety.

"The late-entry log for this crate," Aisen continued, keeping his voice level. "In the past seventy-two hours. I'm asking before the bench is sealed because after the handover I'll need to file a formal request and wait three business days."

"You have observer status only—"

"I know what I have." He kept the clipboard in his peripheral vision. "I'm not asking you to show me confidential materials. Late-entry logs are procedural records. They're part of the crate's chain of custody, which is going to be the first thing the inquiry examines if the Provost calls one. I'm asking to see it now so that I can confirm whether what I'm smelling on this housing is what I think it is, which would make this an inquiry-level situation rather than a procedure failure."

The steward's hands adjusted their grip on the clipboard.

He reached for the ledger on the table.

Aisen had time to read three lines before Master Wei arrived from the left — his approach was always quietly sudden, a quality Aisen associated with people who moved well and didn't announce themselves — and read over his shoulder the same three lines.

The late-entry log showed a single entry: forty-four hours ago, 21:14, an agent from Patron House Arenthal had accessed the crate to request expedited shipment of two reserve crystals. The agent's name was recorded. The releasing steward's signature was also present, which was procedurally correct.

Master Wei's breathing did not change. He had the controlled respiration of someone who had been in difficult rooms before and understood that their response was being observed.

"The signature belongs to the evening duty supervisor," he said quietly. "Who is currently not on shift."

"When does she come on?" Aisen asked.

"Tomorrow morning." He paused. "Unless she is summoned tonight."

Above them, Provost Maren was descending from the gallery with the careful pace of someone whose primary concern was the optics of the moment rather than its substance. Two Patron House representatives were coming down with him, one of whom Aisen recognized as a clerk whose primary function, based on several months of architectural study of who attended which institution at what time, was documentation and report-filing.

They were going to close this room and write the first report before Aisen could establish what the evidence actually showed.

He made a decision.

"The residue on the housing is disruption paste," he said — to Master Wei, quietly, at a volume that did not carry to the Provost. "Third-century formulation. It's in the restricted Archives under the old nomenclature, which means whoever used it either has access to restricted materials or has very specific historical knowledge. Either way, it's not a supply error."

Master Wei looked at him with an expression that was specifically the expression of a man recalculating an estimate he'd already made. "You've read the restricted materials catalogue."

"I have a sponsor arrangement with the head librarian."

"Since when?"

"Three years."

Master Wei said nothing for two seconds. Then he moved to intercept Provost Maren before the Provost reached the bench.

Aisen stayed where he was and did not look directly at the bench, which now had a quality of solved-problem about it that he was trying not to let show on his face. The want he'd walked in with — get to the evidence before it disappears — had been partially achieved. Master Wei now had the same information he had, which meant two people knew this was deliberate, which meant the first-report narrative would encounter at least one competent counter-voice.

That was not a resolution. It was a position.

\emph{From a position you can operate,} he thought, which was a sentence his father had said to him four years ago in a context that had seemed to be about chess and which he now understood had never been about chess.

He waited.

Amara found him twenty minutes later outside the hall, in the archway that led to the procurement offices, where she had apparently been waiting with a patience that was its own form of information-gathering.

"I heard the bench fail," she said. She was not breathless — she moved through the Academy's corridors in the way of someone doing a thousand things simultaneously and seeing all of them clearly, which was how she always moved. "Mnemonic cord."

"Later," he said. "Not here."

She nodded, with the economy of someone who understood exactly what \emph{later} meant in context.

The cold from the hall's stone floor had worked its way through his shoe soles and was in his heels now, a dull ache that tracked up the backs of his calves. He had been standing on stone for an hour and a half. He would need to move, soon, or the ache would become a cramp and a cramp in the examination hall's aftermath would be the kind of visible physical involuntary thing that he had spent five years learning to prevent.

He walked to the window at the end of the archway.

Outside, the Academy's courtyard was occupied by students doing the thing students did after something went wrong during a formal trial: standing in small groups, conducting post-mortems, building the narrative from the scraps of what they'd seen. The student who had been at the bench — Dorin — was at the courtyard's north end with two others, and from this angle Aisen could see the set of his shoulders, which was not the slump of public embarrassment but the controlled rigidity of someone who was angry.

\emph{He knows it wasn't his fault,} Aisen thought. \emph{He knew it the moment the bench failed.}

That was useful information. It meant Dorin had either been prepared for sabotage or had the situational sensitivity to recognize it in real time. Either characterization was worth noting.

He pressed his thumbnail into the inside of his left wrist — not hard, just a pressure point — and held it there until the cold in his heels receded enough to think clearly.

The inquiry would take three days minimum. In three days, the late-entry log would have been reviewed, the evening duty supervisor would have been questioned, and the Patron House Arenthal agent would either have been located or would have become conspicuously absent.

Whichever of those outcomes materialized would tell him which families had considered it worth the risk of sabotaging an Academy examination for the purpose of creating an incident in which Covenant involvement could be publicly justified.

That was a hypothesis, not a conclusion.

He wrote it in the merchant's cipher in the back of the small notebook he kept in his jacket's inner pocket, using the abbreviations that would mean nothing to anyone who didn't have the key, and locked it there.

Then he went to find somewhere to sit down before his heels decided the matter for him.
