% Auto-generated from 11-Academy-Initiation.md
% POV: Aisen | Landscape: City | Location: The Divided City

The smell of the Academy's dormitory block was institutional soap and recent whitewash and the particular minerality of water that had been moving through old stone pipes since before any of the current occupants were born. Aisen noted this in the way he noted all environmental inputs in unfamiliar spaces: automatically, without deciding to.

He had been assigned to a room with two other first-year students. The assignment was evidently alphabetical or random — the room contained, by the time he arrived with his bag at half-past ten in the morning on the day of intake, a student named Theron who had arrived earlier and already made his bed with the precision of someone raised in an orderly household, and a student named Marcus who arrived forty minutes after Aisen and assessed both of them with the brief calibrating glance of someone deciding what they were dealing with.

Aisen unpacked his bag in eleven minutes. He used the smaller bedside cabinet for the items he would access daily and the larger bottom drawer for everything else, and he oriented the bed so that his back would not face the door while sleeping because this was a habit he had developed five years ago and had not been able to break since.

He registered that he was building a cover persona immediately, which was correct. He also registered that the cover persona and his actual self were not as clearly separated as they had been during the city operations, because here he was going to be in the same space with the same people for months, and the operational distance that made city-work clean was not available.

The Academy's first week was a delivery system for context: who the instructors were, what the coursework structure would be, what behavioral norms were expected and which were merely observed-upon. Aisen attended all of it and noted it and said what was approximately useful during the mandatory introduction circuits and shook the hands of the people whose hands were extended at him and asked the questions that were expected without asking the questions he actually wanted answered.

The Headmaster's convocation address on day two used the phrase "contribute productively to society" four times, never specified what society or who determined the form of the contribution, and never used the word \emph{Covenant}. Aisen recorded this in the back column of his notebook under the taxonomy heading \emph{Invisible Mechanisms}, which was his private vocabulary for things that operated by not being named.

On day eight, he found a small piece of paper in his dormitory room that should not have been there.

The Academy's wards were not subtle about their function. They were classroom-grade: calibrated to prevent items from being teleported in or out, to log magical signatures that entered the space. Paper introduced through non-magical means — slipped under the door while the wards' attention was on the magical spectrum — was a gap in the surveillance system that anyone patient enough to identify it could exploit.

The note read: \emph{Library. Third floor. East section. Midnight.}

He held it for thirty seconds. Then he burned it over the room's candle — which was within the range of items students were permitted to burn in rooms, specifically tallow and wax, and a small piece of paper produced insufficient smoke to trigger any monitoring — and he sat on the edge of his bed for another five minutes and thought about probability distributions.

Probability one: this was a genuine network contact, using the gap in the ward coverage that Harald had told him to look for, timed to day eight because that was apparently the agreed window for initial activation.

Probability two: this was a Covenant probe. A test item, delivered by security officers monitoring new students for suspicious behavior. If Aisen appeared at that meeting, he would be confirming a suspicious profile.

He could not determine which was more likely. He did not have enough information to weight the distributions. What he had was the following: Harald had said there was a network inside the Academy. Harald had said he would be contacted. Harald had not said how. A midnight meeting in a library was consistent with what network contact would look like. It was also consistent with what a Covenant probe would look like.

The correct response to this ambiguity was to attend the meeting in such a way that discovering it was a probe produced no actionable information — to go, to be present, and to say nothing that could not be said by innocent curiosity or a false invitation.

At eleven forty-three he left the dormitory.

The movement was easy — the Academy's corridors at this hour were quiet but not monitored for presence, only for suspicious behavioral patterns, and there were enough students who moved at night for study or personal reasons that a single person with a deliberate-but-not-furtive gait produced nothing notable. He moved through the corridor, down the east stairway that he had established his fastest route through on the second day, and into the library's west entrance, which was unlocked but staffed at night by one person at the main desk who was currently reading and did not look up.

The third floor's east section was old materials — archive-tier texts from earlier periods of Covenant history, not restricted but not commonly accessed. He expected the shelving to smell familiar. It did: old paper, cloth binding paste, the faint bitter quality of long-stored ink.

A woman was standing between two shelves with her back to the third-floor window. She was approximately thirty-five. Her hands were empty and at her sides in the specific relaxed posture that trained people adopted in meeting spaces to communicate that they were not immediately dangerous, which was itself a piece of operational vocabulary that told him she'd had some training.

He said nothing. He waited, because he did not know enough yet to speak usefully.

"You came early," she said. "That's one indication."

"Indication of what?" he asked.

"That you're not here by accident." She moved slightly — not toward him, just a postural shift. "I'm Cynthia. I work as a research librarian here. Harald sent information ahead. But I need to establish a few things before we continue."

Aisen noted his own inclination to say \emph{go ahead} and held it. She had named Harald — that could be knowledge obtained from Harald under duress, or it could be genuine. The network's security model assumed compartmentalization even upward. He should not give Cynthia anything until she demonstrated she had information he hadn't given her.

"What did Harald tell you about me?" he asked instead.

She looked at him for a moment with the expression of someone recalibrating. "He said you'd do that," she said. "He said you'd ask the question that required me to demonstrate prior knowledge before committing to the exchange. He said it's a habit that he's had difficulty training out of you and that at this stage he's decided it's more useful than not."

That was specific enough to be either genuine or very well prepared. He waited.

"He said you ran the Verin Lane operation," she continued. "Recruitment disruption. He said you confirmed through Thrace the day before the officers arrived. He said you burned the note before coming here, which is why I had to commit the information to a non-written form."

She paused. "He also said your left hand goes cold after extended Flux use, which you correct with specific physical pressure, which means you've been doing this long enough to develop compensatory habits."

The left-hand cold was something Aisen had never written down. Operational security protocol: physical symptoms were not recorded. Harald had told him this at fourteen and it was still in practice.

He exhaled once. The operational layer settled slightly. He said: "What do you need from me?"

"Nothing except your understanding of the situation," Cynthia said. "I don't need you to run operations. You're not ready for Academy operations yet — you don't have enough local context. What I need you to know is the surveillance map, the contact network, and the communication protocols here so that when you are ready, or when something requires your specific capability, we're not introducing all of this under pressure."

She produced a folded paper from her jacket. A floor plan.

He took it and opened it and looked at it with the full attention it required. Seven marked locations. He noted them spatially in the mental map he had been building since arrival. Cross-indexed against what he'd observed of the surveillance patterns in the first week.

"These are the seven dead zones in the Academy's observation coverage," Cynthia said. "They change slightly each term as equipment is updated, but the east library section" — she indicated where they were standing — "has been consistent for three years. The Covenant knows a perfect surveillance grid is not possible and typically identifies the gaps and monitors the gaps, but this section hasn't been actively monitored long enough that its gap status appears to have drifted into administrative assumption."

"Which means it's either genuinely unwatched or they've decided to let it be unwatched for the purpose of seeing who uses it," Aisen said.

"Yes," Cynthia said. "Both possibilities are functional. If it's genuinely unwatched, we use it. If they're watching to see who uses it, we need to be the kind of people who would plausibly be here."

"A librarian working late. And a student who can't sleep and wanders."

"Both of which are visible and documented and produce no suspicious pattern." She retrieved the floor plan. "The contact network within the Academy: three instructors, five elder students, two administrative staff. I won't give you the full list tonight. You'll receive one contact at a time, with enough context to verify them, and they'll know one piece of information about you that I'll tell you so you can confirm they're genuine."

He stored this process.

"What do you need from the network?" she asked. "Not operationally. Functionally. What kind of support helps you do what you're supposed to be doing here?"

"Access to texts the coursework doesn't assign," he said. "Specifically anything on Covenant administrative history from the primary period, pre-Academy governance models, and the Covenant's internal structure as it existed before the current form." He paused. "And I have a standing arrangement with Soren at the city Archives. If that arrangement becomes inaccessible — if Soren's capacity declines before he makes the liaison notation — I'll need a way to maintain the access he was going to formalize."

Cynthia was quiet for a moment. "Soren's capacity," she said. "I wasn't aware of that timeline."

"He told me approximately three years."

"That's shorter than I understood." A pause. "I'll look into the liaison notation procedure."

She left at twelve twenty-three, moving back through the shelves at the pace of someone who belonged here at this hour. Aisen stayed for seventeen additional minutes and read two pages of an administrative history from the shelf nearest him, which he retained approximately half of, and then walked back to the dormitory by the corridor route that passed the security monitoring station in a directly visible way, so that if he was being observed his behavior would produce a readable narrative: student, insomniac, library visit, returned.

In the room, Theron was asleep. Marcus was also asleep, breathing audibly.

Aisen sat on the edge of his bed and pressed his thumbnail into the inside of his left wrist and held the pressure point until his chest had returned to baseline. The unnamed sensation was there again — a tension locked into his jaw that his private taxonomy filed under \emph{physical register, unresolved category} — and he held the pressure point until it was present but manageable.

He was in the Academy. The network was here. Cynthia appeared to be genuine. Box J through R existed thirty minutes across the city, and its timeline was three years.

He had an enormous amount to do with no clear starting point and no visible clock except the one Harald had given him: \emph{you'll know when the moment exists.}

He lay down. He told himself to sleep. After some time that he did not measure, he did.

\textbf{Writing Notes}:
- Tone: Alert, observant, efficient communication
- Dialogue: Instructions delivered through information exchange
- Imagery: Library as sanctuary; surveillance as architectural design
