% Auto-generated from 02-Training-Years-Begin.md
% POV: Aisen | Landscape: Mountains | Location: The Border Pass

He had not slept.

He knew this because the candle on the table had burned from finger-height to the stub and the wick was drowning in its own pool, and the light coming in under the shutters had shifted from street-lamp amber to the flat grey of early morning without any gap that he could account for. He had been sitting on his bed with his knees drawn up and the wall cold against his back and he had thought, methodically, until there was nothing left to think that he hadn't already thought twice, and then he had thought it again anyway.

\emph{Someone saw me move the cup.}

He had been eleven when it happened the first time — not a cup, a key that had been just out of reach. He'd thought he'd imagined it, the way the key had slid two centimetres across the shelf while he was looking at it. The second time he had been eleven and a half and it had been a cup, exactly as Seraphine had said, and he had told his father that same night because the alternative was carrying it alone and he had not been built for that kind of carrying.

His father had not been surprised.

\emph{That should have told me something.}

The sounds from the kitchen had started at what Aisen estimated as two hours before dawn: the scrape of the spice box being moved aside, which meant Harald was reaching the shelf behind it, which meant the blade. The whetstone's rhythm followed — unhurried, meditative, the sound of someone doing the one physical thing they knew how to do that required no decisions.

Aisen got up. His feet found the floor's boards in the dark without thinking, muscle memory from a thousand early mornings, avoiding the third plank from the wall because that one gave a note that the others did not. He pulled his jacket on over his sleep clothes and went to the doorway.

Harald did not look up. He was at the table with the blade laid across two folded cloths and the whetstone moving in the flat strokes that kept an edge's angle true. The candle near his left hand was fresh — he had lit a new one at some point, which meant he'd been awake long enough to watch the first burn down.

"Sit," his father said.

Aisen sat. The table between them was bare except for the blade and the whetstone and a cup of something that had gone cold some time ago. The kitchen smelled of the bone-soup stock Harald had left simmering when they went out last night, now reduced to a dark thick paste that had scorched slightly at the edges of the pot.

He needed to ask the question, but he needed to ask it correctly.

There were two versions. The first version was: \emph{who is she?} That would get him a prepared answer — a history, a structure, the shape of the Covenant as his father understood it. That answer existed in his father's mind, already organized, already edited for what a twelve-year-old should probably not know yet. He would learn something from it, but he would not learn the \emph{true} thing.

The second version was harder to find.

He watched his father's hands on the whetstone. The stroke was eight inches, heel to tip, consistent pressure through the curve at the end. His father had shown him this motion — not as a lesson, just practically, \emph{this is how you do it so you don't ruin the edge} — and at the time Aisen had thought it was about knives. He understood now it was about learning what consistency felt like in your hands so that when something was inconsistent, your body flagged it before your mind caught up.

\emph{His hand shook last night. The ring finger. He didn't want it to.}

"You knew her before last night," Aisen said.

The whetstone paused for one stroke. Then resumed.

"Not personally," Harald said. "I knew who she was. I knew what she worked for. I knew she had been assigned to the region three years ago." He turned the blade, checked the light along the edge. "I didn't know she knew about you."

"But you knew someone would, eventually."

His father set the whetstone down.

This was the thing that Aisen had been approaching during the hours of sitting on his bed with his back against the cold wall. Not the Covenant as an institution. Not Seraphine as a specific threat. The thing he needed his father to confirm, because he was almost certain it was true and being almost certain was worse than knowing.

"You've been preparing for this since before I moved the key," Aisen said. "You'd already started teaching me before I showed talent. The rules about how to sit, how to move, what not to look at, how to exit a room. That wasn't general carefulness." He kept his voice even. The evenness cost him something he couldn't name, some muscle in his throat that wanted to do something else. "You already knew there was something in me that would eventually need to be hidden."

Harald was quiet for long enough that the wicked wick on the candle sputtered and recovered and sputtered again.

"Yes," he said.

Aisen pressed his thumb against the edge of the table's corner until he felt the dull line of pressure wake up the nerves there. Something to hold onto while he processed.

"Why didn't you tell me?"

"Because children who know they're in danger behave differently than children who know how to be careful." His father picked up the whetstone again but did not use it, just turned it in his fingers. "If I told you at seven that someone might kill you for what you could do, you would have lived in fear. Fear makes people visible. It makes them choose badly in small moments — the wrong hesitation, the wrong response. I needed you to practice caution as a habit, not as a strategy you reached for when you were frightened."

"So you let me be frightened last night instead."

"You were ready for it last night." Harald looked up then, and Aisen could read the tiredness on his face — not a night's absence of sleep, which was accountable, but something older. "You were ready to be frightened and still do the right thing. I've watched you do that in smaller situations for two years. Last night was not a small situation, but the skill transferred."

Aisen turned this over. It was logical. He could see the reasoning. He could also see that the reasoning existed separately from the question of whether it had been fair, and he was old enough to understand that fairness and necessity were different problems that sometimes had the same answer and sometimes didn't.

He filed the fairness question. It wasn't useful right now.

"What happens now?" he asked.

"Now you learn the things you weren't ready for before." Harald set the whetstone on the cloth beside the blade with the careful placement of something he was done with for the morning. "Not tactics. Not the structure of the Covenant — that's information, you can read and memorize information. What you need is harder to teach than information."

"What is it?"

His father stood and moved to the window. He drew the shutter back two centimetres — not to look out, just enough to let the grey morning air in, which was colder than the kitchen air and thinner, carrying grit from the street below and the faint animal smell of the yard. The city was beginning to wake: somewhere below, a cart wheel grated against the gutter of a downspout, a sound that meant someone was moving deliveries before the street traffic made it impossible.

"Judgment," Harald said. "The difference between information about when to act and the actual felt sense of \emph{now}. The difference between knowing that trust must be earned and being able to tell, in a specific person in a specific moment, whether the earning has happened." He turned from the window. "Those are not things I can give you in lectures. They're things you have to learn by making decisions and living with the outcomes."

"Then there will be more tavern visits," Aisen said.

"There will be everything," Harald said. He picked up the cup of cold liquid from the table and drank it anyway, without expression, which was its own form of statement. "Come. I'm going to show you the maps. There are systems in this city that are larger than the Covenant as an organization — older, in some cases — and you need to understand what you're inside before you can understand the part the Covenant plays in it."

He moved to the door of the small room beyond the kitchen, and Aisen followed.

The maps were already on the wall. He had seen them before, or thought he had — but what he had seen before was geography: streets, districts, the river, the guild boundaries marked in different-coloured lines. What he saw now, because his father began pointing and explaining and using the language of systems rather than locations, was something else entirely: the \emph{texture} of control, the way authority moved from one institution to another, the pressure points where disruption could travel fastest and farthest.

He had a thought, brief and not quite formed: \emph{this is what a knife feels like when it's been sharpened correctly.} Not the blade — the feeling in the whetstone, the consistency of the work, what it meant to have prepared something for the use that was coming.

His father had been sharpening him for seven years.

He pressed that thought small and set it aside. There would be a time to feel about it. This was not that time. His father was pointing to the district around the Academy and explaining the function of the placement — geographic, symbolic, practical — and Aisen was twelve years old and this was the first morning of the rest of his life and there was a lot of it left to cover before dark.

He looked at the map.

He began to learn.

\emph{What followed was not swift.} The months compounded into years the way stock compresses in a pot: the volume diminished but the concentration increased, and by the time Aisen was fifteen he carried more situational knowledge about the city's power structures than most adults who worked within them, because most adults who worked within any system were trained to see it from inside, which was the one angle it could not be seen clearly from.

The training was not elegant. It was conversational, iterative, full of reversal: Harald would present a problem, Aisen would work through it, Harald would point out the three things Aisen had not considered, and they would work through it again.

On a Tuesday in autumn, when Aisen was thirteen, Harald dropped a tin cup on the kitchen floor while Aisen's back was turned. By the time Aisen flinched and looked, his father had already moved two paces to the left.

"Where was I standing?" Harald asked.

Aisen pointed to the spot. His father shook his head. "Where was I standing when you \emph{heard} the cup?"

Different question. Aisen closed his eyes and reconstructed the sound — the direction, the floor's resonance under the impact, the half-step he'd heard without registering it. He pointed again, six inches to the right.

"Better," Harald said. "Do you know why that matters?"

"Because if someone's moved by the time I look, I need to know where they were."

"Because most people only track where things \emph{are}. You need to track where things \emph{were}, because the difference is intent."

And Flux.

They never used that word inside the building. Harald called it \emph{the work} in the same flat voice he used for everything that required care. Aisen practiced it at the limits of perception: small movements, brief durations, only in isolated rooms with windows shuttered.

At fourteen, he held a coin suspended for nine seconds and felt the warmth in his sternum close like a fist. He sat on the floor afterward with his back against the cold plaster wall, his left hand trembling, and counted his heartbeat until the count stopped mattering. Harald brought him water without comment and sat across the room reading a logistics manual while Aisen's body rebuilt what the work had spent.

At fifteen, the one time he pushed past the limit, the headache lasted three days — a pressure behind his eye sockets that moved in waves synchronized to his pulse. He lay in the dark with a wet cloth over his eyes and understood, in his body rather than his mind, what his father had meant about maintaining a tool by not breaking it. Some lessons only properly lived in the bones.

The years were not dramatic. The city grew familiar from walking the same streets at different hours, in different weather, reading the same blocks as they changed over seasons. He kept observations in a notebook disguised as a merchant's ledger — figures in columns whose patterns, read against a memorized key, yielded everything.

Harald aged in ways that accumulated in the mornings — slower to straighten from the table, stiffer in his wrists, the skin around his eyes thinning until the veins showed blue. There were nights when Aisen could hear him through the wall: awake, not moving, just existing in the dark with whatever weight he had carried since before Aisen's first memory of him.

Aisen didn't ask. His father had given him the structure for understanding what was coming; asking him to also carry the weight of Aisen's worry was a cost he had no right to levy.

The final meeting came on the morning of Aisen's eighteenth birthday, in the same tavern — the Broken Wheel — where Seraphine had arrived six years before.

Harald had ordered two cups of the house ale. Neither of them drank.

"I can't come with you to the Academy," Harald said. No preparation. No lead-in. The sentence delivered the same way he delivered everything that mattered: flat, factual, already decided. "If they're watching for connections to me, your position is compromised the moment we're seen together."

Aisen had known this was coming. He'd known it for two years, the way you knew weather was changing before the clouds arrived — in the body, in the pressure, in the quality of the air between them that had shifted from teaching to something closer to farewell without either of them naming it.

"I understand the Academy is a Covenant institution," Aisen said. "Designed to identify threats through co-option or removal."

"Yes. But it's also the only place you'll find others like you." Harald's hands were on the table — no whetstone, no blade, nothing to turn. "People who understand that systems are constructed rather than inevitable. Who might work toward something different."

He pushed an envelope across the table. The paper was thick, the seal plain.

"Contacts. Encoded. Study it completely, then destroy it."

Aisen took the envelope. It weighed almost nothing.

"What happens if I fail?" he asked.

"You die," Harald said. "Probably before twenty-five."

"And if I succeed?"

"Then you become responsible for everything that comes after. That's a burden that won't end."

The silence between them held the weight of seven years of mornings at the table with the whetstone and the blade and the maps on the wall and the bone-soup stock scorching at the edges of the pot. Aisen pressed his thumb against the table's corner — the same gesture, the same pressure, the same nerves waking beneath the skin.

Harald stood. "I'm proud of you. And I'm sorry."

"Would you do it differently?"

His father was quiet for three seconds. "No. Because the alternative is leaving you ignorant while systems destroy the people around you. I can carry your resentment. What I couldn't carry is knowing you could have helped and didn't because I failed to prepare you."

He left. The door closed behind him with the hinge-squeak that Aisen had memorized six years ago, the first night, when Seraphine had walked through it and everything had changed.

Aisen sat in the Broken Wheel with an encrypted envelope and a cup of ale he didn't drink and the knowledge that his childhood was over and had been for a long time, and whatever came next would be the consequence of six years of whetstones and maps and small correct decisions made in the dark.

He was eighteen. He was ordinary. He was completely, comprehensively prepared for something that would probably kill him.

He left the tavern. He did not look back.
