% Auto-generated from 20-The-Threshold.md
% POV: Aisen | Landscape: Coast | Location: The New Harbor

The woman's name wasn't in the manifest. She'd been brought in as a reference source — someone observed speaking sympathetically in the market about the occupation — third on the interrogation log for the day. Not even a primary target. A circumstantial secondary whose only crime was saying the soldiers seemed young to the wrong baker at the wrong time.

Aisen was assigned to warehouse security. He stood at the far end of the room, facing the main door, his boots on concrete that smelled of motor grease and the metallic residue of cleaning solution that never quite masked what had been cleaned. His job was to watch the door. He watched the door.

He could hear it. The elevated tone the interrogators used for pressure sequences — not shouting, never shouting, just a pitch designed to compress the space between questions until the silence felt heavier than the words. The cadence of repeated questions. A long silence, and then a sound that had no other word for it — wet and bright and final, like a spark hitting water.

He didn't look. Looking would accomplish nothing and would be noticed.

Twenty-two minutes. She had known nothing — the baker from the north market who'd told someone three weeks ago that the soldiers seemed young — and they'd used full pressure protocol before arriving at that conclusion. When they released her, she walked past Aisen toward the door and he saw, peripherally, the left side of her face: magical burn, subcutaneous, the skin darkened along the jawline in a pattern that followed the nerve paths. Three to four weeks to heal, if she had access to treatment. She carried herself with the specific held-still quality of someone who had learned in the last half-hour that movement had costs.

He opened the door for her. The hinge moved without sound. The air from the corridor was cooler against his face.

She didn't look at him.

He closed it again and returned to his position at the far end of the room. His thumbnail found his inner wrist — count to twelve, steady rate. He counted. The concrete under his boots held the cold the way only institutional floors did: deliberately, as if the building had been designed to ensure that comfort was never accidental.

That evening he encoded a report for the dead-drop: \emph{Interrogation protocol applied to civilian reference source with no resistance connection. Magical contact burn. No intelligence extracted. Pattern of application suggests non-graduated approach — pressure protocol applied at intake regardless of assessed threat level. Recommend provincial networks understand this as tactical information: post-contact civilians will not produce actionable intelligence but will generate visible trauma that functions as local messaging.}

He encoded it in three sections to keep each below detection threshold. It took forty-five minutes. He wasn't sure it mattered. He encoded it anyway.

Ashfall had forty thousand people and a three-battalion Covenant military presence. The Southeastern Territory was the early-window operation — Sector 4-North, in the deployment framework he'd seen in the Academy database. The first of the seven.

The city's administrative district had been converted to command operations: the former council hall holding Colonel Serath's command post, the surrounding blocks organized into supply, communications, and detention infrastructure. The residential quarters around the perimeter were occupied nominally — residents still present, movement still permitted within curfew hours, streets watched at the checkpoint intersections.

He'd been in the territory eleven days when he found the supply crates. Non-standard markings, concentrated density, stored in the back section of warehouse three behind standard military equipment. When he asked a senior soldier about them, the man said: "Administrative. Not our concern." He photographed the marking notation with the index camera and included it in a separate transmission.

Kael arrived in the city in the second week. She was presenting as a displaced merchant, which gave her legitimate reason to visit military administrative offices requesting resource allocation guidance. Their encounter in the supply warehouse — she was requesting information about allocation protocols — lasted four minutes. She told him the provincial resistance was activating: supply chain disruptions, infrastructure, targeted operations against isolated units. Not direct engagement; the Covenant forces were triple the size of anything the regional networks could field directly. But occupation costs could be raised.

"What do you need from me?" he asked.

"Force distribution. Small-group patrol patterns, supply chain nodes, command decision cycle timing. Where are they concentrated and where are they exposed."

He transmitted it in five pieces over three weeks.

The first resistance attack — the convoy ambush, three dead, seven wounded — happened in the nineteenth day of occupation. Military security response was immediate: heightened alert, expanded checkpoints, curfew shortened by two hours, interrogation operations activated.

The interrogation protocols expanded to encompass wider reference categories. The baker from the north market was the forty-third person logged that week.

On the thirty-first day of deployment, Aisen was summoned to Colonel Serath's command post.

He'd been anticipating a different version of this moment for some time — a detention summons, or a directed confrontation from military security. The actual summons was routine in format: his unit commander told him Colonel Serath had requested a meeting, no further information provided, third hour of afternoon.

Micro-want: leave the meeting with his operational cover intact.

Ticking clock: unclear. Which was itself a form of ticking clock.

Serath's command post occupied the former council chamber — high ceiling, stone floor, the carved relief maps of the territory on the walls that the council had used for agricultural planning and that now served a different purpose. Serath was at the central table with two documents in front of her. She was approximately fifty-five, with the specific compact quality of someone who had spent decades making decisions under physical constraint. She indicated the chair across from her.

He sat.

"I received a transmission nine days ago from an instructor at the Central Military Training Facility," Serath said. "The instructor's name is Amara. You participated in her cohort during the training rotation."

He held his face steady. "I did."

"She told me that you'd been gathering intelligence for resistance network contacts throughout the training rotation, and that your intelligence had been reaching provincial networks that are currently operating against occupation forces in multiple sectors." Serath's voice was factual — information delivery, the same register as Amara's countdown in the database room. "She told me that you were doing this because you believed accurate information about military operations would produce better outcomes than information vacuum in provincial resistance structures. And she told me she'd been facilitating part of your training in that direction."

He counted three full seconds before speaking. "Is this a detention situation?"

"It's a conversation," Serath said. "I'll tell you why if you want context."

"Yes."

"Because I've commanded military operations for thirty years in eight different regional territories and I've watched two models play out in repeated sequence. Maximum suppression produces eliminated organizations, significant civilian casualties, regional resentment that lasts two to three generations, and recurring resistance that reconstitutes with different structural form approximately eight years after suppression. Constrained suppression produces disrupted organizations, lower civilian casualties, shorter resentment cycles, and lower probability of reconstitution." She looked at him directly. "Amara believes I'm a commander who prefers the second model. She's not wrong."

He assessed the statement. It was the right framing for someone who wanted to present as a pragmatic rather than ideological actor — which was exactly the framing that would make the subsequent offer seem institutionally legitimate. It was also possibly true.

"She's offering me as an asset," he said.

"She's offering context," Serath said. "For a secondary role. You'd be coordinating intelligence about provincial resistance structure — leadership, resource allocation, organizational vulnerabilities — and you'd be transmitting it to regional command in a form that informs our operational targeting." A pause. "In exchange, I commit to using that targeting intelligence for constrained approaches. Disruption rather than elimination. Leadership isolation rather than casualty maximization."

The sub-sternum sensation coiled tight and held — it had been there since she'd said Amara's name, a warmth that sharpened to a point between his shoulder blades. He held it in parallel with the analysis.

The structure of the offer had a problem. If he accepted, he'd be providing resistance network intelligence to Covenant command. Some of that intelligence — even if used in constrained approaches — would result in network members being detained or killed. The provincial networks he'd been working with for years would be damaged by information he supplied. The people who'd trusted his encoded transmissions would become targets.

The counter-argument was that if he refused, someone without his understanding of constrained versus maximum approaches would fill the intelligence role — someone with no existing resistance network framework, no relationship with Kael's methodology, no reason to curate toward minimum damage. And Serath would conduct operations with whatever intelligence she could gather regardless.

He ran the model twice and it came out the same both times. The dominant variable wasn't whether resistance intelligence would be used — it was whether the person using it understood what to do with it.

"What's the source protection model?" he asked.

Serath looked at him for a moment. "Meaning?"

"If I'm providing intelligence about provincial resistance operations, I need to understand how you protect the sources and methods. Not for my protection. For theirs. If you're extracting information from the network contacts I've been working with, the value of the network degrades every time you extract from it. Source protection isn't ethics — it's operational maintenance."

She was quiet for three seconds. Then she picked up one of the two documents from the table and set it aside. "We would operate on a tiered system. Actionable targeting intelligence would be used once and the access point would be closed. Non-actionable pattern intelligence would feed into predictive modeling without specific operations. In no case would source identifiers be preserved in command records."

He'd need to verify that. He couldn't verify it now.

"I'll need forty-eight hours to assess," he said.

"Good," Serath said. The word carried no warmth. "Hesitation I can work with. Eagerness makes me nervous."

He walked back to his unit barracks through the administrative district. The checkpoint soldiers nodded at him as he passed through. He was wearing his military uniform, carrying military credentials, walking along a military patrol path. He was a person who worked for the institution that had interrogated the baker from the north market.

He pressed his thumbnail to his inner wrist, counted to twelve, and kept walking.

He'd need to encode a transmission to Kael before the forty-eight hours concluded — not to ask permission, because that wasn't the operational framework, but to ensure that Kael's network had updated information about the risk landscape before he made any decision that changed it.

He encoded the transmission that evening in four segments. He left the first segment in the dead-drop point and waited.

He didn't sleep much. He didn't need to.

\emph{Act II ends here. Aisen is 19, deployed in the Southeastern Territory, thirty-one days into provincial military operations. He has been working both military and resistance intelligence roles simultaneously. He now faces the threshold decision the chapter is named for: whether to move from passive intelligence gathering to active double-agent operation within military command.}
