% Auto-generated from Chapter-14-Kaelen-at-Academy.md
% Source: C:\Users\PCGAME\Desktop\story&universes\chain-weapon-story\finished-manuscript\manuscriptbaseforchapters\Chapter-14-Kaelen-at-Academy.md

\chapter{Kaelen at the Academy}
\renewcommand{\CurrentChapterNum}{14}
\POV{\glyphAisen}{Aisen}

Aisen noticed the ivy first.

It had been dead for weeks—a brittle lattice of brown stems clinging to the dormitory's eastern wall, desiccated past any reasonable hope of recovery, the kind of plant the academy groundskeepers left in place because removing it would have required scaffolding and the removal of dead things ranked low on institutional priority lists that dealt in living ones. It had been dead when Aisen arrived. It had been dead through three rainstorms and two lectures on theoretical applications of elemental resonance. It was, by every metric available to someone who had grown up watching things rather than growing them, irreversibly dead.

And now it was not.

A single tendril of green had pushed through the brown—not at the base where new growth would logically begin, but halfway up the wall, at the precise height where someone passing beneath might have brushed it with a shoulder. The green was vivid against the stone, startling in the way that living things are startling when they appear in places that have been surrendered to lifelessness, and it curled toward the morning light with the unhurried certainty of something that had been given permission to resume.

Aisen was still studying it when the boy walked around the corner of the dormitory and stopped.

He was Aisen's age but built differently—lean in the way that suggested constant physical work rather than deliberate training, wiry and sun-darkened, with hands that were visibly calloused even at a distance of ten paces. His hair was brown and tangled with what appeared to be, upon closer inspection, an actual leaf—not decoratively but incidentally, the way debris gathers on someone who has been moving through vegetation rather than past it. His clothes were practical, patched at the elbows and knees in fabric that did not quite match the original, stained with earth in patterns that had become permanent features rather than accidents. Dirt lived beneath his fingernails with the permanence of a resident, not a visitor.

But it was the eyes that held Aisen. Green-tinted—not the pale green of shallow water but the deep, saturated green of light filtered through canopy, the color that exists only in places where living things have stacked themselves so densely that the sunlight must negotiate passage through each layer before reaching the ground. And his skin, where the morning light caught it at certain angles, carried a texture that was not quite skin—a faint roughness, a suggestion of bark, as though the boundary between this boy and the things he tended had been negotiated to a compromise rather than a clean border.

The boy looked at the ivy. Then he looked at Aisen looking at the ivy. His expression was not guilty or surprised but simply present, the way someone looks when they have done something as natural as breathing and are faintly puzzled that it requires explanation.

"It was thirsty," he said.

His voice was quiet and direct, undecorated, the voice of someone who had spent more time talking to things that did not require social calibration than to things that did. Aisen filed the observation alongside the green tendril and the bark-textured skin and the leaf in the hair and arrived at a preliminary assessment that would, over the following weeks, prove almost entirely correct: this was a person for whom the living world was not a subject to be studied but a conversation already in progress, and the academy's stone corridors represented not an opportunity but an interruption.

"Kaelen," the boy said. He extended a hand. His grip was firm and dry and faintly warm, the warmth of someone whose circulation ran closer to the surface than was typical, and Aisen felt in the contact a sensation he could not immediately categorize—a faint hum, sub-auditory, like standing near a beehive or pressing an ear to a tree trunk in wind.

"Aisen Korv."

"I know. Third floor, room seventeen. The window bed." Kaelen's mouth moved in what was technically a smile but was too quick and small to be called one—a flash of expression that appeared and vanished like sunlight through a gap in the canopy. "The groundskeeper told me. She talks to the hedges. The hedges remember everyone who passes."

Aisen waited to see if this was a joke. Kaelen's expression suggested it was not.