% Auto-generated from 31B-Force-Geometry.md
% POV: Serath | Landscape: Wasteland | Location: The Ash Fields

The math said eight of them would die.

Serath stood at the edge of the tree line and counted the variables the way he counted anything — by their spatial relationships, their force vectors, their probabilistic outcomes in the next sixty to ninety seconds. Fourteen fighters in his unit. Twenty-two enemy combatants holding the ridgeline fortification. The fortification was a field-constructed stone-and-earth barrier, approximately chest height, reinforced with Flux warding on the forward face that his scouts estimated at Tier 2 sustained — which meant approximately 6 MU per minute of maintenance cost spread across however many casters they had behind the wall.

The terrain between the tree line and the fortification was open ground. Sixty meters of grass, mildly upslope, with no cover larger than a fallen log at approximately the thirty-meter mark. A killing field. Any approach across that ground, against a defended position with Flux-capable operators, was a math problem with a predetermined answer.

Eight dead. Maybe seven if the approach was fast and the suppression worked. Maybe nine if it didn't.

Captain Roth wanted him to take the position. Command wanted this ridgeline cleared by nightfall because the ridgeline controlled the sight lines for the next three kilometers of the valley push, and the valley push was on a timeline that did not accommodate alternative approaches.

"Serath." Roth was behind him, looking at the same open ground, reaching the same calculations. Roth was forty-one, a career military officer with enough campaigns behind him to know exactly what a sixty-meter killing field cost. "What's your read?"

"Direct assault costs us more than half the unit. Even with suppression, the warding on the wall gives them a three-to-five-second defensive advantage against anyone crossing the open ground. At sprint speed, open ground takes twelve seconds to cross. That's nine seconds of exposed approach against at least four Flux casters."

"Alternatives?"

Serath had spent forty minutes on alternatives while the scouts were counting heads behind the fortification. "Flanking route exists on the eastern side — a ravine that runs parallel to the ridgeline, covered by brush. It terminates sixty meters east of the fortification at a point where the wall is lower because the terrain drops. But the ravine is single-file and takes twenty minutes to traverse, which means we can't use it for the main force. If I take four people through the ravine, we hit the eastern flank at the same time the main force demonstrates against the front."

"Demonstrates. Not assaults."

"Demonstrates. Enough Flux noise from the tree line to make them look forward. My four hit the wall where it's lowest, roll the eastern end, and force them to split their defense. Once they split, the main force crosses the open ground against a reduced front with divided attention."

Roth considered this. "The four who go through the ravine will be in close combat against a fortified position."

"Yes."

"With twenty-two defenders."

"With however many defenders rotate east when we hit the flank. Probably six to eight."

"Four against six to eight. In close quarters. At a wall."

"That's the math."

"And you'd be one of the four."

"I'd lead the four. I need Tessaly for her barriers, Grenn for blade work, and Yael for suppression fire on the transition."

Roth looked at the open ground again. The wind moved the grass. Behind the fortification, a sentry shifted position — visible as a shadow-shape against the skyline, the silhouette of a hooded figure adjusting a monitoring crystal on the wall's parapet.

"What's the math on your four?"

Serath answered precisely. "Tessaly and Grenn are above average in close combat. Yael's suppression capacity will buy us the first exchange. We hit the wall fast, breach the eastern end, and create a gap that the main force can exploit within thirty seconds of engagement. If the main force crosses at that moment, the defenders are fighting on two fronts with a broken wall. I estimate two casualties in the flanking group. Probably three. Probably us."

\emph{Probably us.} Not a calculation he flinched from. Death was spatial — it occupied a specific location in the future, defined by the intersection of force vectors and the body's tolerance for damage. Serath had stood at enough of those intersections to understand them as geometry rather than destiny.

"Go," Roth said.

The ravine was worse than the scouts described.

Not wider and not narrower — the width was accurate, about a meter and a half, passable in single file. The problem was water. The ravine's floor was a drainage channel for the ridgeline above, and three days of rain had left it running with a shallow stream that turned the bottom to slick clay under a hand's width of moving water. Footing was a constant negotiation between speed and stability, and the negotiation favored neither.

Serath led. Tessaly behind him, then Grenn, then Yael at the rear. They moved in silence except for the unavoidable sound of boots in water, which was masked — partially — by the wind moving through the brush above the ravine's lips.

Tessaly was twenty-six, a former stonemason who had transitioned to military service when her village was incorporated into a military district and stonemason work disappeared into requisition contracts. She had 140 MU capacity and specialized in barrier construction — defensive Flux work that created temporary physical obstacles. She was steady, quiet, and had the particular kind of physical patience that came from years of working with stone: you moved when it was time to move, and you didn't move when it wasn't.

Grenn was thirty, a blade specialist with the unhelpful distinction of being the best close-quarters fighter in the unit and the worst long-range operator. His Flux capacity was 70 MU — enough for personal augmentation, not enough for ranged combat. He compensated with physical skill that Serath had watched, over two years, and classified as genuinely exceptional: fast, economical, and controlled in a way that came from talent refined by constant practice rather than natural instinct alone.

Yael was twenty-four and could sustain area-suppression fire — broad Flux pulses that disrupted concentration and degraded defensive wards — for approximately ninety seconds before her 110 MU reserve was depleted. Ninety seconds was the window.

The ravine took twenty-two minutes. Two minutes longer than estimated because of the clay footing. Serath adjusted the timing calculation: if Roth began the frontal demonstration at the agreed signal — a triple Flux pulse visible from the fortification — the main force would need to see the flanking engagement begin within fifteen seconds of the signal to maintain coordinated pressure.

They reached the ravine's terminus. The brush thinned. The fortification's eastern end was visible: a wall section approximately waist height where the terrain dropped, constructed from stacked fieldstone with earth packing. No warding glow on the eastern face — the defenders had concentrated their Flux reinforcement on the forward wall, facing the open ground. Standard allocation error: defend the obvious threat axis, underinvest in the secondary.

Behind the wall, Serath could see three defenders. Two standing, one crouched — the crouched one appeared to be maintaining the warding on the front wall, which meant he was a dedicated caster locked into sustained output. The two standing defenders were blade-equipped sentries scanning the eastern approach. They weren't scanning hard. Their attention was forward, toward the tree line, where the main force was presumably visible.

Three visible. The other seventeen to nineteen were distributed along the wall's hundred-meter length, concentrated on the front. When the flanking attack hit, some would rotate east. How many and how fast would determine who lived.

Serath held up three fingers. Tessaly, Grenn, and Yael acknowledged. He pointed to the wall section, traced the approach route — twenty meters from the ravine terminus, through low brush, to the wall's eastern end — and assigned positions: Grenn first over the wall (blade engagement, close the nearest sentries), Tessaly second (barrier on the wall behind them, seal the eastern section, prevent reinforcement), Yael in the ravine mouth (suppression fire, keep the caster disrupted).

Serath would go with Grenn.

He sent the signal. Three Flux pulses into the air above the ravine — a burst pattern that Roth would see and recognize. The signal was also visible to the defenders, but its origin point — east, behind them — would create two seconds of confusion: \emph{what's behind us?}

Two seconds.

Grenn moved.

Twenty meters of low brush at full sprint took Grenn four seconds. Serath was a half-step behind, three-tenths of a second slower because his stride was shorter. The sentries at the wall heard them at the fifteen-meter mark — the sound of two bodies crashing through brush was not subtle — and both turned.

The nearer sentry reacted well. Weapon up, weight shifting to a defensive stance, focus crystal in the off hand beginning to gather charge. She had perhaps one second before Grenn reached the wall.

Grenn reached the wall.

He went over it the way Serath had seen him go over obstacles in training — a single fluid motion, left hand on the wall top, body rotating over the edge with the blade already in trajectory. The nearer sentry's defensive stance met Grenn's blade at the intersection point between "prepared for attack" and "the attack was faster than prepared for." Grenn's blade took her across the forearm — not the kill, the disarm — and the combat crystal fell from her numbed hand as her flexors failed.

Serath came over the wall one second later. The second sentry — male, older, heavier build — had his weapon oriented and was swinging at Grenn's exposed back. Serath blocked with his own blade, a crossing guard that redirected the sentry's swing upward and away. The redirected blade passed over Grenn's head by six centimeters. Grenn didn't flinch. He'd heard the blade pass and trusted the geometry.

Serath pressed the block into a bind — his blade riding the sentry's from guard position into the bind-and-lever that forced the sentry's weapon out of alignment. The sentry was stronger — twenty kilograms heavier, at least, with the leverage advantage of a longer blade. Strength was the wrong variable. The correct variable was angle. Serath adjusted his binding angle, shifted his weight to his rear foot, and let the sentry's strength push the bind past the tipping point where force became overextension.

The sentry stumbled forward. Serath cut him across the back of his right hand — tendons, not lethal — and the blade dropped.

Two sentries, both disarmed. Three seconds total. Grenn had already moved to the crouching caster, who was still locked into the forward warding — the sustained Flux output meant he couldn't disengage quickly without collapsing the entire front-wall defense. Grenn's blade found his shoulder. The caster broke concentration with a shout, and the warding on the front wall flickered.

From the ravine mouth, Yael's suppression fire began. A broad Flux pulse rolled across the wall from east to west — a wave of disruptive energy that degraded defensive focus, scattered concentration, and created the specific kind of cognitive noise that made combat magic unreliable. Defenders along the wall staggered. The warding flickered again and this time, without the dedicated caster maintaining it, collapsed.

The front wall was unwarded. Sixty meters of open ground was no longer a killing field — it was just open ground.

Five defenders rotated east. They came along the inside of the wall at a run, weapons drawn, reacting to the flank breach with trained speed. Five was within the estimate. Five against three in close quarters (Yael was in the ravine, providing suppression, not available for melee) — Serath, Grenn, and Tessaly.

Tessaly placed her barrier.

It materialized across the wall's interior, a translucent plane of compressed Flux energy that blocked the approach corridor between the eastern wall section and the rest of the fortification. The five incoming defenders hit it. The first one — running too fast to stop — struck the barrier at full speed and rebounded hard enough that Serath heard the impact of his body on the packed earth. The other four stopped short, two of them attempting to breach the barrier with Flux strikes.

Tessaly's barriers were her best work. 20 MU to create, 4 MU per minute to sustain, and constructed with the precise structural understanding of someone who'd spent years building actual walls out of actual stone. The Flux strikes hit the barrier and dissipated. It held.

The barrier bought thirty seconds. In those thirty seconds, the main force crossed the open ground.

Serath watched them come from behind the barrier — Roth leading, ten fighters running across the grass at full sprint, no longer under Flux fire because the forward warding was down and Yael's suppression was degrading the defenders' ability to organize ranged response. The defenders behind the barrier were turning — realizing the flanking attack was a fixing action and the real threat was the ten fighters approaching the front wall.

They were too late. The main force hit the unwarded front wall at speed, and the fortification became a close-quarters engagement on two fronts, which was the geometry Serath had designed.

It took ninety seconds after the main force arrived. Ninety seconds of ugly, close-range fighting along a hundred meters of wall — blade work, short-range Flux exchanges, the violence of two groups of armed people fighting in a confined space with stone walls on either side. Serath fought during those ninety seconds with the specific focused attention that combat required: each exchange a geometric calculation, each movement a vector adjustment, each decision about angle and distance and timing made at the speed his body could execute.

He took a cut on his left forearm — a defender's blade getting through his guard during a transition between opponents. The cut was deep enough to impair grip but not deep enough to sever tendons. He switched his blade to his right hand and continued. Ambidextrous blade work was not his strongest skill, but in close quarters against disoriented opponents, adequate technique was sufficient.

Grenn was magnificent. There was no other accurate word. He moved through the defenders with an economy of motion that Serath, who understood blade work well enough to teach it, could only classify as art. Not flourishes, not dramatic technique — pure functional combat, every strike serving exactly one purpose, every movement covering exactly one threat. Three defenders fell to Grenn in those ninety seconds. All disabling wounds, not kills. Grenn had the control, even at close-quarters speed, to choose where his blade landed.

Tessaly held the barrier for seventy seconds before her maintenance cost exceeded her reserve and the barrier dissolved. By then, it didn't matter — the defenders on the other side of the barrier were now caught between the flanking group and the main force, and their options had reduced to surrender or die.

Fourteen surrendered. Eight were incapacitated by wounds. None were dead.

Serath sat against the inside of the wall after the fortification was secured. His left forearm was wrapped in a field bandage that was already soaking through. The cut throbbed in a rhythm that matched his heartbeat, which was elevated but declining — the specific physiological signature of a body transitioning from combat state to recovery.

Around him, the unit was processing the aftermath. Prisoners secured, wounds treated, the fortification's defensive positions repurposed for their own use. Roth was at the front wall, directing the establishment of a forward observation post.

Casualties. Serath counted.

His flanking group: one significant wound (his forearm). One minor wound (Grenn had a shallow cut on his thigh that he'd acquired at some point and not mentioned). Tessaly and Yael, uninjured.

Main force: three significant wounds — a broken collarbone from a Flux concussion impact, a blade wound through the left hand, and a deep laceration across the back that would require treatment Serath's field medic couldn't provide here. Two minor wounds that were already bandaged and ambulatory.

Total: five wounded, zero dead.

The math had said eight dead. The math, specifically, had said eight dead for a direct frontal assault across sixty meters of open ground against a warded fortification. The flanking approach — the ravine, the twenty-two minutes of bad footing, the four-person breach team hitting the eastern wall — had changed the math.

Geometry. Not heroism, not inspiration, not tactical genius. The simple, physical fact that sixty meters of approach against a warded wall was a different equation than twenty meters of approach against an unwarded wall, and the difference between those two equations was the difference between eight dead and zero dead.

Serath looked at his arm. The blood through the bandage was dark red, venous — no arterial involvement. He'd have full use of the hand in ten days, reduced grip strength for three weeks, permanent sensitivity to cold weather at the scar site.

Bodies were structures, like fortifications. They had load-bearing elements and non-load-bearing elements, and the difference between a survivable wound and a crippling wound was which elements the blade found. His blade had found forearms and hands and shoulders — the non-load-bearing elements of a human being's capacity to fight. Grenn's blade had found the same targets. The defenders they'd disabled would heal. Their capacity to fight would return.

This mattered to Serath, though he had difficulty articulating why. He had killed before. He would likely kill again. But the effort to preserve life where combat geometry permitted preservation was not sentimentality — it was professional precision. A stonemason who could build a wall did not knock down walls he wasn't directed to. A surgeon who could cut precisely did not cut imprecisely. A fighter who could disable did not kill when disabling was available.

The line between available and unavailable was narrow. Serath walked it with the attention it required.

Roth approached. "Ridgeline secure. I've messaged Command. They want the liaison officer's assessment for the valley push planning."

The liaison officer. The Academy kid with the analytical gaze and the chain-weapon. Serath had worked with him twice — brief interactions, professional, the specific kind of exchange that happened between a combat operator and a liaison officer who understood that his role was to observe and report rather than to direct. The Academy kid was smart in the way that young analysts were smart: comprehensively, systematically, without the experiential calibration that turned intelligence into judgment.

He'd develop judgment. Or the war would develop it for him. The process was the same.

"Send him the terrain assessment," Serath said. "Include the defensive warding configuration we found — Tier 2 sustained on the forward wall, nothing on the flanks. If the next position uses the same doctrine, the flanking approach works again. If they've adapted, we'll need a different geometry."

Roth nodded. "Your arm."

"Functional in ten days."

"I need you functional in five."

Serath looked at the ridgeline. Below, the valley extended east — green, forested, bisected by the river that the campaign would follow toward the provincial interior. Somewhere in that valley, the next fortification waited, built by people who were defending their territory with the resources they had against a military force that outnumbered and outequipped them.

The geometry of the next engagement would be different. Every engagement's geometry was different. That was the fundamental truth of combat that no training simulation captured: the math was never the same twice.

"I'll be functional in five," Serath said.

Roth left. Serath sat against the wall, left arm held against his chest, and began calculating the approach to the next position. The terrain maps were in the liaison officer's assessment pack. He'd review them tonight.

The wind moved across the ridgeline. Inside the fortification, the prisoners sat in a group under guard — fourteen people who had woken this morning as defenders and would sleep tonight as prisoners. The transition between those two states had taken ninety seconds and one flanking maneuver through a ravine.

Geometry. The shape of force applied to space. The difference between who lived and who died.

Serath understood this. It was the only thing he understood completely.
