% Auto-generated from 15-The-Heresy-Archive.md
% POV: Aisen | Landscape: Dungeon | Location: The Deep Archives

The wards shifted when he pressed the crystalline authentication against the lock plate — a flutter at attention's edge, less a sound than a pressure change, the same half-felt quality as preservation magic in Section 7 but quieter, working on different frequencies. Then the sealed door opened inward without a sound.

Section 6.

Aisen stepped through and let the door close behind him.

Eleven minutes in. He had — at most — forty before the checkpoint monitor cycled its status records and noticed the temporary credentials active in a section they weren't issued for. He'd calculated the window twice on the walk here and arrived at the same number both times: forty minutes, probably less, enough only if he didn't stop to read.

The micro-camera Cynthia had pressed into his palm that morning was flat and cool against his right index finger, secured under a strip of adhesive above the first knuckle. He held his hand slightly curled, natural-looking at a glance, the lens uncovered.

Section 6 was smaller than Section 7. The first thing he noticed was the smell — older than Section 7, drier, the preservation compound here not the sharp mineral-and-cold of active research magic but something duller, a smell like old paper sealed in darkness for decades without disturbance. No amber indicator lights anywhere. No evidence of recent reconfiguration. Whatever was happening here was about movement, not research: the shelving units were standard century-old Archive construction, unchanged, and the disorder on every third shelf — gaps where materials had been pulled, inventory notations taped to the empty spaces, half-filled transfer crates stacked in the aisle — was recent.

He moved to the first major section and began photographing.

Military records. Bindings coded in standard Archive notation, the format Cynthia's guide had taught him to read in the three days between receiving it and entering this room — the color of the binding thread, the placement of classification marks on the spine, the weight and cut of the cover cloth. He recognized them correctly: \emph{Previous Covenant Military Operations – Regional}. \emph{Enemy Force Assessment Frameworks – Third through Fifth Centuries}. \emph{Conflict Historical Analysis – Suppression Operations}. Three of the four transfer crates were already filled with these. Whoever was moving them had prioritized military history above everything else.

Photograph. Move. Don't slow down to read the text, just document what's present and let the analysts translate it.

The second section was smaller. Research materials, more recent bindings, tighter classification marks. \emph{Covenant Combat Advantages – Comparative Historical}. \emph{Magical Suppression Techniques – Practical Implementation}. \emph{Weapons Development – Institutionally Classified}. His hand tracked the edges of a bound volume before he'd decided to touch it — an involuntary movement, fingers reading the cover texture while his attention was already three rows ahead. He withdrew the hand.

Document only. Don't move the materials. He photographed the section from two angles and moved on.

It was the third section that stopped him.

Not a decision to stop. His feet stopped, and the rest of him caught up.

The materials here didn't use standard Archive notation. The bindings were pale — some kind of treated skin rather than cloth — and aged in a way that was different from the dust-and-dry-paper age of the military records. These had been preserved under heavy magical maintenance for a very long time and were old underneath the preservation. The organizational codes on the spines were a system he didn't recognize: not numeric, not alphabetic, a combination of small pictographic marks that the guide in his interior pocket (he'd committed most of it to memory, but the physical copy was there) hadn't covered.

The padded containers they were being moved into were separate from the transfer crates: smaller, lined with what looked like a dampening material, the kind used for objects that couldn't tolerate stress without degradation. Someone was treating these materials as fragile in a way that military history documents weren't.

There were perhaps forty volumes. He photographed them methodically — spine, cover where visible, the pictographic organizational marks, the padded containers open beside them. What he couldn't read, the analysts would. What he couldn't photograph without moving the materials, he left.

The sub-sternum sensation was there: not quite discomfort, not quite anticipation, something without a clean name that he'd learned to treat as a register of significance rather than danger. Harald had used to say that a good intelligence framework should treat anomalous response as data even when the source wasn't yet identified. The sensation wasn't fear. It was something more like: \emph{this is the part that matters}.

He didn't know why these materials were being moved with more care than records of military campaigns. He knew that he didn't know that, and he photographed the entire section twice.

Twenty-six minutes in.

He was photographing the third row of paleograph-marked spines when he heard the footsteps.

They came from the adjacent corridor, not from the Section 7 direction — from what his mental map of this level placed as the administrative office. One person, unhurried rhythm, not the maintenance worker's shuffle-and-pause but the steady walk of someone crossing from one point to another with a destination in mind.

Aisen moved to the nearest shelving unit and positioned himself sideways in the aisle, one hand on a transfer crate's edge, facing the shelves. His breathing was even. The movement had been unhurried and deliberate — any observer glancing through the section entrance would see someone standing at a shelf making notes, not someone who had just frozen.

He counted footsteps. Seven, eight, nine — the sound grew, then didn't. The administrative worker had turned into the aisle south of Section 6, parallel but not entering. The footsteps continued and faded toward the stairwell at the far end of the Archive level.

Aisen waited two full minutes before he moved again.

Thirty-four minutes in. He needed four minutes minimum to reach the exit corridor, pass back through Section 7, ascend to the checkpoint, and reach Corvin's desk with the appearance of someone who had simply lingered over research materials too long. He photographed the last three visible spines in the archaic section and stepped back.

There was one padded container that had been sealed — the dampening cloth cinched at the top with a cord. He didn't touch it. Cynthia had been specific: document, don't handle. And beyond the specific instruction was the operational principle he'd internalized from years of working with Harald's framework: a sealed container in a restricted section wasn't evidence of what was inside, only of the fact that someone had decided its contents required this much protection. That was itself the finding.

He photographed the sealed container from three angles.

Then he moved to the Section 6 door, waited for the ward to re-engage behind him, and walked the corridor back toward Section 7 at the pace of someone who had found what they were looking for and was satisfied.

Corvin was still at the checkpoint desk, a cold cup of something to his left and a bound ledger open that he was not reading. He looked up at Aisen's approach with the particular unseeing glance of someone who had been alone for three hours.

"Finding what you needed?" he said.

"Reference materials." Aisen set his research clearance card on the desk in the same motion he used every week. "Ended up going farther into the stacks than I expected. Thanks for waiting."

"That's what the desk is for," Corvin said.

Aisen climbed the stairs at his standard pace. At the top he crossed the Academy's ground-level corridor without stopping, exited through the east courtyard, and kept walking until the Archive entrance was out of sight. Then he pressed his thumbnail to the inner wrist, counted twelve seconds, and let the count reset his breathing.

The camera passed to Cynthia the following morning through an encounter in the bookseller's row near the south Academy gate — a narrow miss on a wet footpath, Aisen catching her satchel strap, her hand closing around what he pressed into her palm during the steadying contact. She didn't look at him after.

Two days later, Caris gave him the analysis.

The military records were as expected: historical pattern documentation for Covenant military interventions in regional territories where resistance movements had appeared. The magical suppression research was consistent with what Section 7 had suggested — active technical preparation for operations against organized magical practitioners. Together, Caris said, the two categories pointed toward a coming military campaign aimed not at territorial expansion but at organizational elimination.

"Provincial territory," he said. "They've identified organized magical resistance at scale. The suppression precedents they're reviewing show the pattern: military force to eliminate leadership structures, then institutional reorganization to prevent reformation."

"The archaic materials," Aisen said.

Caris was quiet for a moment. "That's what we didn't expect."

Their analysts had spent a day on the pictographic organizational system before identifying it. Pre-Covenant era — more precisely, from a specific period three centuries prior when the Covenant's institutional authority had still been contested. The materials were documentation of heretical movements: magical groups that had operated outside Covenant doctrinal authority, some of them suppressed violently, some documented only in these archived records.

"Why would they archive records of movements they suppressed?" he asked.

"Because they needed to understand what those movements were doing," Caris said. "If you're going to eliminate something, you need to know how it functioned, how it organized, what made it resilient. Those records aren't just historical. They're operational source material."

A cold pressure settled at the base of his skull, distinctly.

"And they're reviewing those records now," Aisen said.

"Moving them to active storage," Caris confirmed. "Yes."

He brought the analysis to Kael through a second-level courier with a note that used the encoded shorthand she had established for escalation alerts. She called a meeting for the next evening through the same channel.

She was already inside the storage building when he arrived, standing near the east wall, not pacing. He'd noticed before that she rarely paced — deliberate physical economy, the same quality of minimizing unnecessary movement that he'd started to recognize as characteristic of people who had spent years working in spaces where being noticed carried real cost.

"The heresy archive," she said when he finished. She hadn't asked questions during the report: she'd listened, and now she summarized. "They're pulling documentation on how every pre-Covenant magical resistance movement organized and how it was eliminated."

"That's the current analysis."

"Then the provincial networks need to know," she said. "Not that there's a military campaign coming — we already knew that from the conscription data. But that the campaign has historical targeting intelligence behind it. That whoever designs the operation will know which organizational structures to strike to prevent reconstitution." She paused. "How long has the Covenant had access to those archaic records?"

"We don't know. They've been in Section 6 for at least a generation — that's the depth of Cynthia's sourcing inside the Archive. The movement into active storage is recent."

Kael looked at the floor to her left for a moment — not at anything specific, the look of someone accessing an internal model. "The provincial networks don't build the way pre-Covenant heretical movements built. Different geography, different communication methods, different doctrinal framing. But there are organizational parallels they'll recognize." She looked up. "We need military doctrine, not just historical patterns."

"Academy institutional systems," Aisen said. He'd considered this for the past day, since Caris had given him the analysis. It was where the logic pointed.

"Yes. Current combat doctrine, deployment planning frameworks, regional force allocation systems. If we can access those, we can tell the provincial networks what they'll actually face in the field — not what the Covenant did in previous centuries but what they're trained to do now."

"That's a different class of operation," Aisen said. "Archive infiltration with temporary credentials is recoverable if detected. Database infiltration leaves trace patterns that Covenant security actively monitors."

"I know what I'm asking," Kael said.

He knew she did. He had also already worked out the core of the calculation, which was that if the provincial networks encountered a well-planned military suppression without tactical doctrine intelligence, the cost would be paid in specific ways he could model: leadership structures targeted in the first week, communication channels cut, reconstitution attempts failing because they'd repeat the exact organizational forms the Covenant's historical intelligence had identified as vulnerable. People who hadn't been warned would make decisions that got them killed because the decisions were reasonable without context and fatal with it.

"I need a few days to assess access methodology," he said.

Vex listened to all of it before she said anything. Not a courtesy she always gave — she was capable of cutting to the question before he'd finished — but tonight she sat with her hands folded on the library table and let him work through the full account. Caris's analysis. Kael's request. The calculation.

When he was done, she said: "The people who get stopped by this operation — do they know the risk they're running?"

"Some of them. Some of them know the broad shape but not that the Covenant has heresy archive intelligence behind the targeting."

"And if you're caught pulling military doctrine from Academy databases."

"Then I'm detained," he said. "The network loses an asset. Kael's people lose an information source. The provincial networks face the campaign without doctrine intelligence." He kept his voice level. "I've modeled the second scenario in about six different ways. The outcome is consistent."

Vex unfolded her hands and pressed one flat against the table. "I'm not saying don't do it. I'm saying: make sure the risk is yours to take and not someone else's."

"The extraction protocols would put Cynthia at risk," he said. "I know that."

"Does she?"

"She built them." He paused. "She knows."

Vex nodded once. Her expression didn't shift toward reassurance — she didn't do that, had never done that — and he was glad of the absence. What she gave him instead was a specific query about the access methodology, whether timing the database pull against Academy administrative record cycles would reduce the detection window, which it would. They worked through the technical approach for forty minutes, neither of them naming what they were actually doing, which was building the plan.

When the library closed at ten, he walked back to the dormitory through the east colonnade. The stone underfoot had the slight unevenness of the older gallery — this section of the Academy hadn't been resurfaced in his time here. He noticed the texture without trying to.

The sealed container in Section 6, tied with cord, not opened. He turned the image over once and filed it: closed unknown, source insufficient.

But it had been moved with particular care.

He went to sleep eleven minutes after reaching his room, which was sufficient.
