% Auto-generated from Chapter-09-Tarovin-Formal-Apprenticeship.md
% Source: C:\Users\PCGAME\Desktop\story&universes\chain-weapon-story\finished-manuscript\manuscriptbaseforchapters\Chapter-09-Tarovin-Formal-Apprenticeship.md

\chapter{Tarovin's Formal Apprenticeship}
\renewcommand{\CurrentChapterNum}{9}
\POV{\glyphAisen}{Aisen}

Spring arrived in the border town like a debt being repaid—slowly, with reluctance, and only after the earth had been convinced it would not be cheated again. The frost gave ground by inches. Meltwater carved new conversations into the ditches along the north road, and the air carried that particular mineral sweetness that meant the river was swollen and the soil was remembering how to be soft. Aisen noticed these things the way he noticed most things now: involuntarily, catalogued, stored against some future need he could not yet name.

He was thirteen and had begun to understand that noticing was not the same as knowing. The distinction mattered in the same way that watching a river and swimming in it were not the same skill.

Tarovin arrived on the second day of thaw, walking the north road with a satchel that sagged with the weight of papers and the particular exhaustion of a man who had been arguing with institutions. His robe was different from the one Aisen remembered—heavier, with a scholar's pin at the collar that caught the watery light—but his gait was the same: measured, deliberate, as if each step were a sentence in a paragraph he was still composing. His hands trembled faintly at his sides, the way they always had. Aisen had once thought it was cold. He understood now that it was something else: the cost of holding precision in a body that wanted to be imprecise.

Tarovin smiled when he saw Aisen in the doorway of the tavern. It was the patient smile of someone who kept promises to future selves.

"You're taller," Tarovin said.

"You're grayer," Aisen replied, and immediately felt the small heat of having been too familiar. But Tarovin laughed—a sound like a book being closed gently—and set his satchel on the bar as if returning to a desk he had never quite left.

He produced a letter. It was folded into a careful square, the creases sharp enough to suggest it had been folded and refolded many times in consideration. The seal on the outside was the academy's: a glyph of interlocking circles that meant, Aisen had learned from one of his small notebooks, \emph{knowledge through application}.

"This is for you," Tarovin said. "Not to open now. To keep."

"What is it?"

"A recommendation. When the time comes, it will speak for you."

Aisen took the letter and felt the weight of it—not heavy, but dense in the way that intentions are dense. He wanted to open it immediately and had to press that impulse flat like a crease in linen. "What does it say?"

Tarovin's eyes crinkled. "It says you notice things. That matters more than any born advantage."

He said this with the plain certainty of someone stating arithmetic, and Aisen felt the words settle into a place in his chest that had been waiting for them without knowing it. Not pride, exactly. Something more structural—a beam being laid across a gap he had been measuring for years.