% Auto-generated from 06-First-Contact.md
% POV: Aisen | Landscape: City | Location: The Market Ward

The reading room that Soren used for private meetings was not a room Aisen had been in before. It was separated from the main Archives by a door that required a bone token to open — the token was a small flat chip of yellow-white that felt like compressed chalk between the fingers and left a faint smell of cold wax on the skin. Soren had placed Aisen's token on the table beside the library check-in register two weeks ago without explanation, and Aisen had carried it since then in the inner pocket of his jacket, which was the only pocket that closed with a button.

Harald was already inside when Aisen arrived.

The room was approximately four metres by five. Three walls held shelving — older materials, box-stacked, unlabeled, some of the boxes water-stained at the corners in ways that suggested they had been moved from somewhere that flooded before they came to rest here. The fourth wall had two narrow windows, the kind that admitted light on a directional schedule and were currently providing a flat white illumination that made every surface appear slightly colder than it was. The floor was stone. Aisen's feet registered this through the thin soles of his training shoes within the first ten seconds of standing.

There was a woman he had not seen before.

She was approximately forty. Not approximately forty in the round-number sense — she had the specific preserved quality of someone who had spent decades in environments that were cool and dry and low-light, which had the effect of making her look somewhat out of sequence with the age her hands suggested. Her hands were scarred at the knuckle-lines of three fingers on the left in the way of someone who had worked repeatedly with a sharp tool and had accepted the recurring cost of minor cuts in exchange for the precision of the grip.

She was looking at Aisen.

He kept his face still and looked at the table, because the table had something on it.

The map was approximately the size of two open books laid side-by-side. It had been folded at least six times — the fold-lines were deep, some worn to near-translucency — and was now laid flat but not fully flat; the years-of-folding tension kept the corners slightly raised. It was a city map. He recognized the river-bend on the left edge, the market district's characteristic cross-hatch pattern, the bridge the Covenant had widened six years ago which, in this map, appeared as its pre-widening form. The map was at least eight years old.

What was on the map beyond the base geography were marks: small circles, each with a two-letter code and a date. Some circles in red. Some in black. Some in a green that had faded to grey-yellow with age. He counted seven red circles before Harald said, quietly, "Aisen."

He looked up.

"Lyssa," Harald said. "This is Aisen."

Lyssa did not extend her hand. She kept them both on the table beside the map — not touching it, but proximate in a way that made the ownership clear. "How long were you looking at the map before I asked you to stop?" she asked.

"I didn't stop," Aisen said. "You haven't asked yet."

Harald's breathing changed by approximately two percent. Not visible — audible, if you were listening for it. Aisen understood this meant he had said something unexpected, and he ran the sentence back looking for the error, and did not find one, so he stayed with it.

Lyssa looked at him for three seconds. Then she moved her hands back from the table's edge in a motion that was either permission or assessment. Possibly both.

"Then keep looking," she said. "Tell me what you notice."

He noticed:

Seven red circles, six with dates in the past three years, one with no date. Twelve black circles, dates ranging across a span that looked like roughly twelve years from first to most recent. Four grey-green circles with no dates and no codes — those had been marked with a different hand, thicker strokes, the ink quality different from the rest. Three locations he recognized: the Broken Wheel's block, which carried a black circle dated six years ago; the market district warehouse district near the east bridge, which had a red circle dated four months ago; and what appeared to be a building near the Academy's outer administrative district that he had observed from the street but never entered, marked with a black circle and a code that read "DL".

His feet were cold. He moved his weight slightly forward to get more of the ball of his foot in contact with the ground, which helped marginally.

"Forty-one total marks," he said. "Seven red, twelve black, four green-grey, the rest mixed. The red ones are more recent — most in the last three years. The green ones are older and were made by a different person. There are three locations I recognize." He paused. "DL near the Academy administration district has a black circle. That means it's been active, not just surveilled."

Lyssa looked at Harald.

"He counted all forty-one," she said.

"He would," Harald said. "He counts everything."

"It's not counting," Aisen said. "It's that the number ends up in my head before I've decided to put it there." He looked at the map again. "What is DL?"

"A contact," Lyssa said. "Not someone you'll meet until you're in the Academy."

"So the contact is inside the Academy."

"Adjacent to it."

The temperature differential between the insole of his left shoe and his right was now noticeable — his left foot was colder, which meant the floor was slightly uneven and he was distributing more of his weight right. He adjusted.

"The six red circles with recent dates," he said. "Are those locations that were compromised, or locations that were established recently and are therefore new enough to be red?"

Lyssa was quiet for a moment that had the weight of an evaluation. "Both," she said. "Some red means new. Some red means lost." She reached across and covered four of the seven with her palm — not hiding them, just marking them. "These four were lost in the past eighteen months."

He looked at the three she hadn't covered.

"Those three are either new or still secure," he said.

"Yes."

"And you're telling me this because you want me to understand the shape of the thing without telling me where the secure locations are specifically."

Harald's breathing adjusted again.

"He does this constantly," Harald said to Lyssa. "He notices the structure before the content."

"It's the same thing," Aisen said. He wasn't sure if that was rude. He filed the possibility that it was, and continued. "If the four lost locations clustered geographically, that would tell you the Covenant is working from a particular investigative approach. If they're scattered, it means they have information from inside the network rather than surveillance from outside."

He looked at the four covered locations. Lyssa's hand was still over them.

"You don't have to show me," he said. "I'm asking because the answer changes the security model. If it's infiltration, then the problem isn't location security, it's contact security. Which means the DL contact should be compartmented from anyone who was directly connected to the lost nodes."

Lyssa removed her hand from the map.

The four lost locations were: the Broken Wheel block (surprising — he'd had no reading on that), the warehouse district red circle, a location in the southern residential district he didn't recognize, and a small mark near the northern entrance to the city that was far enough from the Academy that it felt like a separate operational zone entirely.

Not geographically clustered.

He kept his face still and did not show that this worried him.

"You'll enter the Academy in four years," Lyssa said. "When you do, you'll be one of approximately twenty people we're aware of in that environment who understand that this network exists. You won't necessarily know who the others are. They may not know you. That's deliberate."

"How do I find them if I need them?"

"You don't find them. You give them the opportunity to find you by behaving consistently enough that someone watching would recognize what they're looking at." She began folding the map with the careful methodical folds of someone who had folded it many times before and understood the sequence. "Your job for the next four years is not to expand your knowledge of this network. It's to not be arrested before you enter the Academy."

"That's a low bar," Aisen said.

"Everyone starts with a low bar," she said. "You'd be surprised how many people fail to clear it."

She folded the last quarter and the map disappeared back into her jacket's inner pocket. The table was empty. The room felt smaller, briefly, in the way rooms felt smaller when the object that had been providing the organizing scale was gone.

Harald was watching him.

Aisen realized he should say something, and then realized he had nothing to say that was more useful than silence, and chose silence. Lyssa nodded once — a motion that didn't invite interpretation — and left through the bone-token door at a pace that was calibrated to be unremarkable.

Harald and Aisen stayed in the reading room for another forty minutes, reading materials that had nothing to do with networks or resistance or Covenant surveillance strategies. Aisen read a volume on the administrative history of the city's guild registry procedures, which was among the driest materials the Archives held and was therefore the ideal cover for being in a restricted reading room at all. He absorbed approximately sixty percent of what he read. The other forty percent was occupied by the map.

Specifically by the four lost locations.

Specifically by the fact that the Broken Wheel block was among them, which meant either the Broken Wheel itself was compromised — which would have operational implications for anyone who had ever used it — or that the circle on the map had referred to a secondary location on that block that he had no knowledge of, which was more likely and also meant there were elements of the network's infrastructure he'd assumed he understood that he had not understood at all.

Walking home through the late-afternoon street noise, the cold still in his heels from the stone floor, he thought: \emph{You've been operating with a partial map.}

That wasn't new information, exactly. But prior to forty-five minutes ago, he'd believed the parts of the map he had were accurate.

Now he wasn't sure which parts were.

He kept walking. The thought locked itself into the same internal ledger where he kept all the things he didn't have enough information to resolve, which was called, in his own private taxonomy, \emph{active unknowns} — distinguished from \emph{closed unknowns} (things he had concluded he would never be able to know) and \emph{solved} (things that had resolved themselves either through information or through events making the question irrelevant).

The \emph{active unknowns} column had been shorter, yesterday.
