% Auto-generated from Chapter-15-Rins-Secret-Mission.md
% Source: C:\Users\PCGAME\Desktop\story&universes\chain-weapon-story\finished-manuscript\manuscriptbaseforchapters\Chapter-15-Rins-Secret-Mission.md

\chapter{Rin's Secret Mission}
\renewcommand{\CurrentChapterNum}{15}
\POV{\glyphAisen}{Aisen}

Aisen first noticed her because she noticed the exits.

Everyone in the dining hall performed their own version of arrival—the noble students with their unhurried drift toward the window tables, the merchants' children with their slightly too-eager selection of seats that signaled aspiration without declaring it, the commoners with their compressed efficiency, sitting where the architecture told them they belonged and beginning to eat before the bread had finished settling on their plates. Everyone performed arrival. Almost no one performed reconnaissance.

The girl at the far end of the center aisle did both simultaneously, and the simultaneity was what caught him.

She moved through the hall the way someone moves through a space they are memorizing rather than occupying—her pace even, her posture unremarkable, her tray balanced with the absent competence of a person whose hands were instruments calibrated for tasks more demanding than carrying food. But her eyes. Aisen watched her eyes travel the room in a sequence he recognized because he performed his own version of it every time he entered a new space: doorways first, then occupants, then the distances between occupants and the exits those distances would become relevant to if the room's social equilibrium shifted from conversation to crisis. She mapped the dining hall in the time it took to cross it, and she did not look at her food until she sat down, and the place she chose—third table from the wall, center-left, a position that offered sight lines to both the main entrance and the service corridor without requiring her to sit near enough to either to be caught in a bottleneck—was not a choice born of social instinct but of operational training.

Aisen set down his fork and watched.

She was his age, or near enough that the difference didn't matter and far enough that it might. Slight—not fragile, but compact, the build of someone whose physical training had prioritized precision over power, every movement trimmed of excess the way a message is trimmed of words that dilute its meaning. Her hair was dark and worn short enough to stay out of her face without effort, and through it ran a single strand of white that she had tucked behind her left ear with a casualness that was itself a kind of concealment—the gesture said \emph{this is nothing, this is unremarkable} in the precise way that unremarkable things do not need to say. Her skin, where the dining hall's amber lamplight caught it along the line of her jaw and the backs of her hands, carried markings that Aisen could not resolve at this distance. Patterns. Faint, pale against pale, visible only when the light cooperated, and it cooperated rarely, as though the markings themselves had been designed for conditions more specific than dining halls.

He catalogued what he could. Filed the rest for later.

Dameris was talking—Dameris was always talking, his commentary a continuous weather system that no silence survived—about the bread's southern provenance and the likelihood that the kitchen staff were cutting the flour with something cheaper, and Aisen made the appropriate responses at the appropriate intervals while the majority of his attention remained fixed on the girl three tables away, who had begun eating with the same deliberate pacing he had observed in Lysandra the previous day: the fork a metronome, the meal a cover story, the real operation something else entirely.

She did not look at him. Not once. And the precision of not looking—the way her gaze swept the room in arcs that passed just over his shoulder or just beneath his sight line, the systematic exclusion of his position from a survey that otherwise covered every occupied table—told him more than looking would have. She knew he was watching. She was managing being watched.

The realization settled into his chest like a stone dropped into still water, and the ripples it produced rearranged several things he had been thinking about the academy's population: that the room contained at least two people running the same kind of operation, and that one of them had better tradecraft than he did.