% Auto-generated from Chapter-37-The-Academy-Purge.md
% Source: C:\Users\PCGAME\Desktop\story&universes\chain-weapon-story\finished-manuscript\manuscriptbaseforchapters\Chapter-37-The-Academy-Purge.md

\chapter{The Academy Purge}
\renewcommand{\CurrentChapterNum}{37}
\POV{\glyphAisen}{Aisen}

The message reached Aisen at the third bell of morning, carried by a boy whose name he did not know and whose face he would not remember—one of Amara's runners, a child recruited from the dockside warrens where orphans learned early that survival was a profession and anonymity its primary tool. The boy handed him a strip of cloth with three words inked in Amara's cramped cipher, waited for no response, and disappeared back into the pre-dawn streets with the practised invisibility of someone who had never, in his short life, had the luxury of being noticed.

Three words. Aisen decoded them against the key he kept in his head—a key that changed every seven days, that he memorised and destroyed, that existed nowhere except in the architecture of his own recall because written keys could be found and found keys could be read and read keys could unravel entire networks in the time it took a clerk to copy them into a report.

\emph{Academy. Dawn. All.}

He was moving before the decryption was complete. Not running—running attracted attention and attention was fatal—but walking with the focused economy of someone who understood that the difference between arrival and absence was measured in minutes and that minutes, in the current climate, were the denomination in which lives were traded.