% Auto-generated from 27-Academy-Conflict.md
% POV: Aisen | Landscape: City | Location: The Divided City

The message from Vex arrived in the dead-drop on day ninety-nine.

He read it at a supply depot collection detail that gave him four minutes of private access to the chalk-notation wall. The notation format was the compressed variant she'd used once before, which he'd learned to read at half-speed: each term indexed to the shared vocabulary they'd built across five years of coded communication.

Translated: \emph{Academy intake suspended. Second-year cohort deployment-relabeling. Regulatory basis: combat deployment authorization for educational institutions. I'm still staff-category, not student-category. Not immediately affected. Vex.}

He set the notation in his working memory and did not re-read it.

He understood what it meant. The Academy had shifted from a two-track system — four-year education for most, early deployment track for a subset — to something different: second-year students were being relabeled as deployable cadets, which under the regulatory provision Vex had identified meant they could be assigned to active provincial operations rather than completing their education cycle.

The second-year cohort would be roughly seventeen-year-olds.

He thought about what he had been at seventeen. He'd been in two years of the Academy's early-track accelerated program at that point. He'd been running an intelligence channel for eight months. He'd been present at the first observation that made him understand what the Covenant's archive was for.

He'd been functional. He was also aware, looking back from day ninety-nine of provincial deployment, that he'd been working from a deficit of specific kinds of knowledge that only the deployment context could provide.

Seventeen-year-olds deployed to province-level operations that involved binding programs and Fourth Circle suppression templates and supply-route ambushes at misassessed junctions was a different thing than what he'd done.

He completed the supply detail.

He asked Amara about it on day one hundred and three, during a technical exchange about the flange-tip fabrication notes.

"The Academy intake suspension," he said. "Second-year cohort deployment-relabeling."

She looked at him with a quality that said she'd been expecting this subject but not from him. "Where did you encounter that information?"

"Personnel communication," he said. "I have contacts still at the Academy."

She set down the fabrication notes. "The regulatory basis is technically valid but operationally premature," she said. "The second-year cohort hasn't completed the formation program that prepares them for provincial assignment complexity. The command rationale is troop density: the occupied territories require more personnel than the standard deployment cycle produces."

"My contacts at the Academy are staff-category," he said. "Not immediately at deployment risk. But attached to second-year instructional tracks."

She understood the implication. If the second-year cohort was deployed, the academic staff supporting them would be reallocated as embedded advisors. Staff-category personnel in that role were deployable under the same authorization.

"I'm aware that this creates a timeline concern," she said. "I'm filing a formal analysis of the second-year cohort's formation readiness gap with the regulatory review board. Not opposition — the regulatory basis is valid. A documented assessment of the capabilities gap and the operational risk of deploying undertrained personnel."

"Will that analysis reach the Registry oversight officer you mentioned?" he said.

She was quiet for a moment. "Possibly. She receives all formal regulatory analysis in the provincial operations category."

He thought about the architecture. He had one channel that could, in theory, transmit to the Registry oversight officer through Amara's formal institutional process. That was an entirely different channel than his resistance-network transmission system. And its function was to provide a counter-argument to the deployment relabeling from inside legitimate institutional process.

He couldn't use it to protect Vex specifically. Targeted individual protection through institutional channels was not the mechanism that was available.

But he could give Amara's analysis more evidence.

"The formation program for counter-insurgency operational complexity requires approximately four months of focused training beyond standard Academy curriculum," he said. "My current provincial operational experience includes at least six scenario types that require judgment capacity which only comes from that training. I can document those scenarios with technical specifics that would support a capability-gap assessment."

Amara looked at him. "Why would you do that?"

"I have a contact at the Academy who teaches second-year magics theory," he said. "If the relabeling moves forward and she's embedded as an advisor in provincial operations, her training and functional preparation are below the threshold for this environment."

This was more direct than he normally operated. He was giving Amara a specific context that was identifiable — not Vex's name, but "second-year magics theory instructor" was a narrow enough category that Amara could determine it if she wanted to. He was choosing the visible risk over the invisible one.

"Document the scenarios," Amara said. "Full technical detail. I'll include them as evidentiary case records in the formation-readiness analysis."

He spent the remainder of the day writing four scenario documentation records. He was specific about operational mechanics and vague about personnel details. He described: the supply-route junction spatial assessment problem; the binding program's intersection with network behavioral patterns; the Serath intelligence exchange as an unnamed "structural analysis requirement"; and the wrap-tip fabrication process as an applied-problem-solving demonstration.

None of the four mentioned his intelligence activity. All four were accurate representations of tasks he'd encountered in deployment that required formation beyond standard Academy preparation.

He gave them to Amara on day one hundred and four.

She read them the same day.

On day one hundred and seven, she told him: "The analysis has been submitted to the regulatory review board. It's entered formal review. I have no way to predict timeline."

He acknowledged this.

He transmitted to Vex that evening through the standard dead-drop: \emph{Formal regulatory review submitted on second-year cohort relabeling. Amara-channel. Timeline unknown. Your formation records — are they documented before or after the relabeling authorization?}

Response came three days later: \emph{Before. Admin recorded my staff classification before the new authorization language. Legal ambiguity exists. Advisor suggests waiting for formal clarification.}

Legal ambiguity: Vex's staff classification had been recorded before the authorization language change, which meant the relabeling might not apply to her under the specific timing provision. This was not security, but it was a delay.

He did not add her to the operational planning in his head. He held the open-unknown column.

On day one hundred and nine, the second-year cohort relabeling was formally implemented at the Academy.

He did not learn this from Vex. He learned it from the personnel supplement that was circulated to all provincial deployment personnel as an operational context update: \emph{Academy intake to resume modified formation cycle. Second-year cohort deployment-eligible effective immediately. Embedded advisory track for qualifying staff.}

"Qualifying staff" was the new categorization. Vex's ambiguity window had closed.

He did not transmit to Vex for four days. This was deliberate: the surveillance pressure on Academy-connected communication would be highest in the first week of the relabeling implementation. He maintained transmission silence on all channels for forty-eight hours except the Orev technical exchange, which continued normally as a legitimate professional communication.

On day one hundred and thirteen, Vex's message in the dead-drop: \emph{Staff-category qualifying determination pending administrative review. Currently in-post. Will advise status. Stay regular.}

"Stay regular" was her signal meaning: don't adjust your pattern for me. She was telling him not to let concern for her situation create observable pattern-disruption in his behavior.

He acknowledged this.

He sat in the barracks common room that evening and participated in a card game for forty minutes. He lost four hands and won one. He did not appear preoccupied. He went to sleep at a standard time.

He was not not preoccupied. But he had learned — through the Aldwyn meeting, through the Serath arrangement, through forty-eight hours of maintaining normal behavior the night after his first transmission — that preoccupation that didn't show was functionally equivalent to not being preoccupied, for every relevant operational purpose.

He thought about what Vex was doing at the Academy tonight: probably working, probably documenting her staff-category determination case, probably teaching whatever classes were still running under the modified formation cycle with the same careful attention to detail she'd always brought to the archive methodology she'd been developing since second year.

She had been building a specific kind of knowledge since before either of them understood what it was for.

He pressed his thumbnail to his wrist for twelve seconds.

Then he stopped. He slept.
