% Auto-generated from 01-Tavern-Opening.md
% POV: Aisen | Landscape: Mountains | Location: The Northern Highlands

The \emph{Broken Wheel} smelled of tallow candles, wet grain sacks, and the specific sourness of ale that had been breathed out by too many mouths for too many hours. Aisen had been sitting in the corner for forty minutes and he could feel the bench's warp in the back of his left thigh—a narrow ridge of wood that had been pressing through his trousers long enough to leave an imprint he'd still feel tomorrow.

He was twelve. Old enough to catalogue the room. Not old enough to do anything useful with what he catalogued.

Forty minutes was the rule. That was how long Harald let him observe before they left—long enough to learn something, not so long that someone learned something about them. Aisen was counting. There were four exits: the front door, a kitchen passage behind the bar, a second-floor staircase, and a shuttered window near the back wall with a bent latch that one hard pull would open. He had found the window in the first three minutes. He hadn't told his father because Harald already knew and the point was for Aisen to find it himself.

\emph{Thirty-eight minutes gone. Two to go.}

That was the want pressing at the inside of his chest: walk out through the front door in two minutes without anyone in this room being able to say, later, that they had noticed him.

He lifted his cup. The water tasted of iron from the well pipe and faintly of the dishcloth that had rinsed the cup last. He let a single swallow sit on the back of his tongue, then set the cup down with no sound, no flourish. His father had shown him how to set things down without announcing them. The lesson had taken a week to feel natural.

Across the room, Mira—the barkeeper's daughter, sixteen and already fluent in the careful art of not hearing what she wasn't supposed to hear—threaded between trestle tables with a tray balanced on her forearm. Two merchants were arguing about a shipment. One of them grabbed at her sleeve. She turned the movement into a half-step away without looking at him, which was not the response of someone who hadn't practiced it. Aisen noted the way her weight shifted to her outside foot: she was always already leaving.

He did not let his eyes linger. Looking too long was the same as pointing.

"Stop staring," Harald said. He hadn't looked up from his drink.

Aisen dropped his gaze to the table. "I wasn't."

"You were deciding whether to feel sorry for her." His father's voice was low, built for rooms like this—it went nowhere except where it was aimed. "She doesn't need that from you. She needs customers who pay their tab and don't grab."

\emph{One minute forty.}

The door opened.

It was the sound that came first: the creak of the Broken Wheel's front hinge, which the owner had calibrated so precisely that the pitch changed depending on how hard you pushed it. A slow open meant someone trying to be quiet. A fast open meant someone who didn't care who heard them. This was neither. This was the door opening at a pace that said: \emph{I know about the hinge. I'm telling you I know.}

Aisen did not look up immediately. He looked at his father.

Harald's right hand moved three centimetres to the left on the table surface and stopped. It was the movement of a man who had started to reach for something and decided not to. His drink didn't move. His posture didn't change. But the tendons in the back of his hand had shifted under the skin in a way that made Aisen's stomach tighten with a recognition he could not name.

He had never seen his father's hand do that.

\emph{One minute.}

Aisen looked at the door.

The woman who had entered was not looking for a table. She was also not looking around the room the way people looked around when they wanted to find someone — that scan, left to right, that declared intention to anyone watching. Instead she had walked eight steps into the tavern and stopped at a point where her back was not to any wall, where she could see the bar, the main floor, the stairs, and the kitchen passage simultaneously. She stood there as if she had simply arrived at where she intended to be.

Her coat was wool, dark, rubbed thin at the elbows. The wear was years old, not recent. Her boots had no mud on them despite the packed-earth street outside, which meant she'd come from pavement, or she had changed them before entering, which was a different kind of attention to detail entirely. Her hair was pulled back in a way that kept it off her neck — practical, not fashionable. Her hands were at her sides with the fingers slightly curled, not fists, not relaxed, but the middle position: available.

Her eyes caught the lamplight.

That was not the right word, Aisen thought. Caught. Her eyes did not catch the light the way glass or water did. They held it, suspended it, processed it into something else. The irises were pale — grey or silver, it was hard to tell in the uneven tallow glow — but their quality was structural. Each time the lamp above the bar swayed in a draught, her pupils adjusted a fraction of a second before the light actually changed, as if her body was receiving information slightly ahead of the event.

Aisen looked away.

His heart was doing something specific: not racing, not lurching, but beating at a rhythm that was two beats per minute faster than usual, which was what fear looked like when it was being kept in order. He had catalogued that sensation before. He associated it with situations from which exit was not guaranteed.

"We're leaving," Harald said.

Not now. We're leaving. The phrasing of a fact, not a decision.

"But we've got—"

"Now." His father's hand found Aisen's shoulder and was already standing, already moving. \emph{He's been tracking the time too. He was past the two-minute mark and he knew it and he didn't tell me because he was watching the door.} "Don't look at her again."

Aisen didn't. He filed that instruction in the same place he filed all of his father's instructions: not as rules but as conclusions someone had reached from experience he didn't have yet. \emph{There will be a reason. Find it later.}

They moved at the pace that didn't cause chairs to scrape or conversations to pause. Harald's hand on his shoulder was a fixed point — neither guiding nor dragging, just present, the way a knot in a rope is present. Aisen kept his eyes at a middle distance, the unfocused gaze of a tired child being taken home, and counted: four steps to the bar's edge, six to the junction of tables, three to the door.

He didn't look back.

He looked at the bartender.

The bartender was already filling a glass he hadn't been asked to fill, at a speed he reserved for people he did not want to keep waiting, set in a spot on the bar he'd cleared with one motion of his elbow. His eyes moved once to the door — the front door, where Aisen and Harald were walking — and then moved away, at the careful velocity of someone who had decided, quite firmly, that he was not going to be a witness to whatever came next.

\emph{She didn't ask for a drink. He's pouring it because he knows what she orders. He's known her long enough to be afraid of making her wait.}

Aisen was thinking this as Harald's shoulder pressed the door open from below the latch — a technique that kept the creak to its quietest register — and then the night air came in against Aisen's face and it was colder than he'd expected. Not cold the way afternoon had been cold, which was even and constant. This cold came from the direction of the river, damp and with a mineral edge — iron in the water, silt and old stone — the kind of cold that crept through wool and settled into the joint where Aisen's collarbone met his shoulder and ached there like a word he hadn't found yet.

Rain was coming. He could smell it the way you smelled things that were still a mile away: not the rain itself but the absence of everything the rain would push out.

They were twenty steps down the street when he heard the footsteps behind them.

Not running. Not following, exactly. Following would have been faster, would have closed the distance. These steps were parallel — same pace, same direction — the way you walked alongside someone to tell them that leaving wasn't actually an option.

His father had already stopped turning before Aisen's body registered what he was doing. The turn was slow, entirely controlled, both hands visible at his sides. Aisen copied the posture because it was clearly correct and he didn't have a better one.

She had covered twenty steps in what should have been eight seconds. She had not hurried. She was not breathing harder than she had been inside.

"Harald." She looked at his father the way you looked at something you had come back to find after a long absence — checking whether the condition had changed, running a quick inventory. "You know what I want to say."

"I know what you're going to say." His father's voice was flat, calibrated. Aisen had never heard that exact flatness before. It was the flatness of a man who had been afraid for a long time and had finished being surprised by it. "That's different."

"Fair." She looked at Aisen. Not an assessment — she had already assessed him the moment she identified him inside, which was probably the moment she had entered. This was something else. She was checking something she'd already measured. "He's twelve."

"Yes."

"He knows?"

"He knows enough."

She made a small sound. Not quite a laugh. The sound someone made when an answer was technically accurate but missed the point. "In my experience, that's not the same as knowing enough."

The cold had found the inside of Aisen's left wrist now, working through his cuff where it had come untucked. He pressed it against his hip without thinking. He was aware, somewhere beneath the careful attention he was paying to every syllable in this conversation, that he was frightened. Not of her specifically — he was doing the arithmetic of that — but of the fact that his father had stopped, which meant his father had calculated that walking would make this worse.

His father never stopped unless stopping was safer.

"You don't have an offer," Harald said. "You have a warning and you've dressed it like an offer."

"The Covenant doesn't make offers to children." She said it without inflection, which was worse than if she'd said it coldly. "The warning is this: he has been noticed. The notice is recent. Whoever noticed him will share that information within the month, because that is what people do with information about children who move objects without touching them."

The air in Aisen's lungs went still.

He had moved a cup. Twice, maybe three times, without intending to — not far, just a centimetre, a twitch, the way a sleeping limb sometimes jerked. He had told his father. He had told no one else.

\emph{Someone saw.}

"Tell him the shape of what's coming," the woman said, "so he can make a smaller target of himself. That's the courtesy. The rest is his arithmetic to do." She glanced at Aisen one more time — the glance of someone who had already gathered what they needed but was filing a final detail. "He's already doing it."

She turned and walked back toward the Broken Wheel's door.

Aisen watched the way her coat moved. It didn't swing at the hem the way coats swung on people who were walking casually. It tracked, close, controlled.

The door opened and settled with its two-pitch creak. Both pitches this time. She'd pushed it at the medium pressure — the one that said: \emph{I know about this, too. I am simply not hiding that I know.}

The door closed.

His father's hand found his shoulder again, lighter now. The shaking was small — only in the last knuckle of the ring finger, a tremor so fine you'd miss it if you weren't paying the kind of attention you didn't need until you needed it desperately.

"Come on," Harald said. "We go home."

"Who—"

"Home first. Questions after."

They walked. The rain had closed the last mile and was beginning, not in drops but in a fine pervasive mist that wasn't quite water yet — the stage before water, when the air had given up trying to hold what it had been carrying. It dampened the sound of the cobblestones under their feet. It made the lamp-glow ahead of them diffuse and directionless.

Aisen's mind did what it always did when he could not speak: sorted.

The woman had known his father's name. His father had known hers — he hadn't said it, but there'd been recognition in the way he held himself when she stopped them, the way you held yourself near something whose weight you already knew. The bartender had been afraid of making her wait. She had walked at a pace that said she had no reason to hurry because nothing he or his father did would change the outcome of the conversation she had decided to have. The Covenant. She had said it as a proper noun, the way you said \emph{city guard} or \emph{tax collector} — a category of thing rather than a name, which meant she assumed they already knew what it was.

\emph{She's right. I don't know enough.}

His father's shoulders had said everything his father's voice had refused to. The hunch that deepened with each block they walked wasn't fear exactly — it was the posture of a man carrying weight he'd been carrying long enough that he'd stopped noticing it most days, but tonight it had gotten heavier.

\emph{He's been expecting this. Not her specifically. Something like her.}

By the time the familiar smell of their building reached him — the floor wax the landlord used, the particular damp of the stairwell that never fully dried in autumn — Aisen had reached three conclusions he was confident in and two he wasn't.

The confident ones: he had been seen using Flux. He was now on a list somewhere. Someone was going to share that list.

The uncertain ones: whether being on the list was the danger, or whether being taken off it was.

He filed the uncertainty. There would be more information. He would find it or it would find him.

He thought, briefly and without ceremony, that this was the last night of something. He couldn't name what, exactly. Just the last night of the version of his life where the things he didn't know were merely gaps, rather than holes with edges sharp enough to cut him if he moved toward them wrong.

He was twelve years old. His collarbone still ached from the cold.

He did not say any of this to his father, because his father already knew it, and saying it aloud would not make it more true — only more present.

He opened the door to their rooms and went in out of the rain.
