% Auto-generated from 20B-The-Two-Masks.md
% POV: Elara Valorin | Landscape: City | Location: The Fortress Hall

Morning began with her father's table.

It was not called a court breakfast. It was called the Northern House Morning Assembly, which was its formal designation in the ceremonial calendar and which conveyed, accurately, that it was not optional. The table was in the east salon of the northern wing, and it seated fourteen: her father at the head, her two older brothers along the right, three advisors and two ambassadorial delegates on the left, and then various members by proximity to relevance. Elara sat fifth from the head on the right, between her brother Calder and the junior secretary to the Agricultural Compact liaison. She had been fifth from the head for three years, which was appropriate for her position, which was the Northern House's third child and only daughter and therefore present for visibility and absent from the actual decision chain.

She wore the grey-blue court dress with the high collar. She wore the silver ear-rings that were her mother's and therefore symbolically load-bearing — they communicated \emph{continuity} and \emph{tradition} and \emph{this family has roots.} She wore her hair pinned with a clip that the court jeweler had made for her sixteenth birthday in a style that was just complicated enough to look deliberate and just simple enough to not require a maidservant, because requiring a maidservant for your hair meant arriving later and she always arrived first.

She touched the crescent scar on her left shoulder through the fabric of the dress, once, before sitting down. It was a habit. The scar did not hurt. It had not hurt in years. The touch was calibration — a reminder of a context in which she had been a different version of herself, in which she had been cold and scared and bleeding and someone had covered the wound with his coat while they waited for a healer. She used the reminder the same way Kaelen used his three knocks on the clay rim, though she had never told anyone this.

"Good morning," she said, in the register that the court called composed and her brother Calder called glassy. She positioned her water glass two centimeters to the right of its place setting, which was the agreed signal.

The signal meant: courier tonight, east garden, third bench.

The system had taken six months to build and was now so embedded in normal social behavior that she could transmit and receive at a formal breakfast table while discussing the Northern House's position on grain pricing with the Agricultural Compact liaison and giving her father's advisors no visible reason to look twice. The water glass was for \emph{meeting confirmed}. An untouched bread roll on the side plate would have meant \emph{meeting postponed, security concern.} A spilled but immediately righted water glass would have meant \emph{abort, surveil me, assume compromised.}

Today was a confirmed meeting.

She ate breakfast attentively. The conversation was about the Compact's new tariff proposals and what they meant for the northern districts' supply chains. Her father was managing the conversation with his usual diplomatic care — appearing to consider multiple positions while communicating that the Northern House's position was already established and had been for weeks. She offered two comments: one factual, one conciliatory toward the delegates. Both were accurate. Both were meaningless. They were the kinds of comments that a useful-but-non-threatening third child made at a table where she was not supposed to be influential.

She was, in fact, very influential. This was information she kept in the same locked compartment as the courier meeting.

"Elara," her father said, at the end of the main part of the meal, using the tone that meant he was transitioning from public business to family business in an official-family way. "There is a delegate this evening from the Covenant Educational Compact. He will want to meet the family formally. Seven o'clock, reception room."

"Of course," she said.

She noted: \emph{Covenant delegate. Seven o'clock.} She filed it next to: \emph{courier meeting, east garden, estimated nine o'clock.}

She would need two hours between them.

The message drop was a simple physical exchange: she walked the east garden path at the route she walked twice a week for exercise, and at the third bench she adjusted her left glove, and during the adjustment she palmed the folded message and left it under the bench slat in a weathered crack she had identified eight months ago. The courier — a woman named Sela who worked in the Academy's library reference section and was one of four couriers Elara had built into the network over eighteen months — would collect it on her own regular garden circuit, an hour later.

Today's message was short and precise. She had encoded it the previous evening in the cipher the network used: a running-keyword system based on a shared text that both parties carried in different formats, resistant to casual interception, not resistant to a professional cryptanalyst with three weeks to work on it. She had been meaning to upgrade the cipher for four months. She had not found a window that felt safe enough for the coordination required.

The message said, in plaintext terms: \emph{Regional assessment attached — Ch18 deployment route confirmed civilian status unresolved. Recommend network-level inquiry Districts 5-6. Contact via Amara channel if military access needed.}

The Amara channel. She allowed herself a moment to think about that — Amara, who had filed a deployment hold on entirely legitimate grounds and who had ridden four hours alone to verify what no one else had bothered checking. Elara found Amara's method clarifying. Amara did not move around the institutional problem; she went directly through it, with documentation, with personal risk, and with the specific confidence of someone who had calculated that the alternative to acting was worse than acting. It was a different style from Elara's, which was to move through the institution invisibly, using its own channels, leaving as few traces as possible. Different styles. She had observed that the network worked better when it had both.

She adjusted her glove. Palmed the message. Set it under the bench slat. Continued walking.

The Covenant delegate was a man named Aldric Fenn, which was not his name in the three other regions where he represented the Covenant, but it was the name on the credentials he had presented to the Northern House protocol office, and the credentials were good enough that the protocol office had accepted them. She knew this because she had a contact in the protocol office who sent her a summary of incoming credentials for high-level meetings. She had not told this contact what to look for; she had told this contact that she was doing a research project on diplomatic procedure, which was true in the sense that she was doing research about a specific diplomatically-linked procedure she needed to understand.

Aldric Fenn looked like someone had assembled a delegate out of correct components: height, grey at the temples, good wool coat, a handshake that was calibrated to convey respect without deference. He was accompanied by two people who were introduced as his assistants and who she assessed, in the twelve seconds she had to observe them before the formal reception protocol required her to look elsewhere, as not assistants.

The reception was forty minutes. Her father made the right sounds. Her brothers were present and visible. She was present, visible, and occupied the ambient social space that third children occupied: warm, non-threatening, slightly ornamental, no one's priority.

Aldric Fenn spoke to her twice. Once at introduction, once during the cup-and-conversation portion. The introduction was unremarkable — conventional, good memory for names, the usual assessment protocol that good diplomats ran on everyone they met. The cup-and-conversation was different.

"The Northern House has been particularly active in what I understand is called the regional assessment process," he said. Conversationally. With a cup in his hand. Looking at the portrait of her great-grandfather on the far wall.

"The Northern House takes its regional obligations seriously," she said.

"As all houses should." He paused. A diplomatic pause, precisely weighted. "Some houses interpret their obligations somewhat broadly. As you know, the Covenant places considerable value on institutional coherence. The definition of \emph{regional obligation} can become -- elastic -- when individual interpretation is allowed too much latitude."

He was still looking at the portrait.

She was still looking at nothing in particular.

"The Covenant's definition of institutional coherence has historically been an important stabilizing framework," she said. Her voice did not have anything in it that a competent observer would call hostile. It also did not have anything in it that a competent observer would call agreeable. She had spent three years developing this register. "The Northern House is committed to frameworks that produce stability."

"Quite." He looked at her then. A brief, direct, assessing look. "I'm glad we understand each other."

She touched the inside of her right wrist — a gesture she had used since she was eleven, when she was running courier routes in disguise and needed something to do with her hands that looked like nothing. An ordinary fidget. A habit. A way of marking: \emph{I have seen you. I have recorded you. You will be in my notes tonight.}

"It is always clarifying to meet colleagues who appreciate precision," she said.

He moved on to speak to her brother Calder.

She wrote the notes at 11 PM, in the cipher, in the notebook she kept in the lining of her second court bag — the one she took to official functions, because taking it to official functions made it a normal item that traveled with her and no one searched normal items. She noted: the Covenant delegate's approach, the specific phrasing (\emph{elastic, institutional coherence, misplaced loyalties among the northern houses} — he had used that phrase with one of the other delegates, she had caught it from four meters away), the two assistants who were not assistants, and her assessment that this was a direct communication, not a fishing expedition. He knew about the regional assessment network. He did not know exactly what she did in it, or he would not have delivered a warning — he would have delivered a consequence.

She had a window. Windows closed.

She wrote a second note, shorter, in a different cipher she used for high-priority items only: \emph{Covenant contact re: network — time-sensitive — recommend accelerated Ch18 follow-up, direct rather than relay.} She sealed it and put the sealed note in the hidden pocket of her third belt — the one she wore under the court dress, the one that had taken her four separate visits to three different craftspeople to commission in components so that no single artisan knew what they were building.

The sealed note was for the courier who could reach the person she trusted to follow up on accelerated timelines without requiring explanation.

She had been very careful about who that was.

She touched the crescent scar on her shoulder.

She put out the lamp.

\emph{— Chapter 20B ends. Camera stays with Elara through final lamp-out. Handoff to Aisen's POV resumes Chapter 21 — the sealed token from Elara's courier chain arrives in Aisen's contact network as an unmarked message. He recognizes the cipher register — he has seen it once before. The crescent marker on the wax seal is new.}
