% Auto-generated from Chapter-36-Covenant-Counter-Strike.md
% Source: C:\Users\PCGAME\Desktop\story&universes\chain-weapon-story\finished-manuscript\manuscriptbaseforchapters\Chapter-36-Covenant-Counter-Strike.md

\chapter{Covenant Counter-Strike}
\renewcommand{\CurrentChapterNum}{36}
\POV{\glyphAisen}{Aisen}

The first reports arrived before the ink on the broadsheets had dried.

Aisen was in the basement of a chandler's shop on the eastern edge of the Trades Quarter—not Tholen's cellar, which had been abandoned three days after the Archive Revelation went public, but a different one, smaller, damper, rented under a false name from a woman who asked no questions because the answers would endanger her children. He was transcribing copied routes onto strips of linen that could be swallowed if necessary when the door opened and Nella came down the stairs with a face that had been emptied of everything except the information it needed to deliver.

"Greyveil House," she said. "Raided at fourth bell. Everyone inside."

He set the pen down. The ink bled into the linen where the nib rested, a small dark stain spreading outward with the slow certainty of something that could not be recalled. Greyveil House was a waystation on the northern corridor—six beds, a hearth, a woman named Dara who cooked lentil soup that tasted of nothing but kept people alive and moving. Dara had three informants who reported to her and a deaf son who carried messages because no one suspected a child who could not hear of listening to secrets.

"How many taken?"

"All of them. Dara, the boy, two runners who were sleeping there after the Velder Street drop went wrong. The neighbours said soldiers. Not city watch—soldiers in Covenant grey, with seal-stamped warrants."

Seal-stamped. That meant the orders had been written before the raids, the warrants prepared in advance, the names already on lists that had been compiled not in the heat of reaction but in the cool administrative silence of offices where men who would never see the people they condemned sat at desks and made decisions with the same mechanical precision that governed tax assessments and grain requisitions. The machinery of state violence did not rage. It processed.

Aisen picked up the pen. Finished the transcription. Rolled the linen strip and sealed it with wax and handed it to Nella, who tucked it into the lining of her belt with fingers that did not tremble because trembling was a luxury that the current situation could not afford.

"The southern corridor?"

"Still open, as of second bell. But Torren hasn't sent confirmation from the mill, and he's never late."

Torren was never late. That sentence, spoken in Nella's flat delivery, contained a grief that neither of them had time to examine. Torren was an old man who mended fishing nets and carried messages in the hollow handles of the tools he transported between riverside workshops, and the fact that he had not sent confirmation meant either that the southern corridor had been compromised or that Torren himself had been taken, and in either case the lentil soup and the deaf boy and the old man's careful hands all joined the same column in a ledger that Aisen was keeping in his head—a ledger of costs that grew longer each hour while the column of gains remained, for the moment, empty.