% Auto-generated from Chapter-23-The-Mysterious-Professor.md
% Source: C:\Users\PCGAME\Desktop\story&universes\chain-weapon-story\finished-manuscript\manuscriptbaseforchapters\Chapter-23-The-Mysterious-Professor.md

\chapter{The Mysterious Professor}
\renewcommand{\CurrentChapterNum}{23}
\POV{\glyphAisen}{Aisen}

Professor Edrik Thane taught modern ritual theory on Tuesdays and Thursdays in the third-floor lecture hall of the academy's northern wing, and he was, by every metric the institution used to evaluate its faculty, unremarkable.

This was the first thing Aisen noticed about him—not an inconsistency but the conspicuous quality of its absence. In a faculty populated by specialists who wore their disciplines like plumage—Maren with her clinical precision, Vellan with his controlled theatricality, the combat instructors with their scarred knuckles and predatory awareness—Thane occupied a register so temperate, so precisely calibrated to the median, that he passed through the academy's social architecture the way water passes through sand: touching everything, adhering to nothing, leaving behind only the faintest evidence that he had been there. He was forty-seven or fifty-two or somewhere in between, the kind of age that eludes precise estimation not because the face gives contradictory signals but because it gives too few signals of any kind—well-maintained, cleanly shaven, the brown hair trimmed to a length that was neither fashionable nor unfashionable but simply appropriate, the institutional equivalent of a room kept to temperature. He wore the faculty's dark wool coat without modification, the collar neither raised against cold nor loosened for comfort but resting as it had been designed to rest, as though the garment had come to him already fitted to the contours of someone who needed no adjustments.

Aisen had been watching him for eleven days.

The notebook recorded the initial observation on the facing page of Naia's entry—not because the two were connected but because the notebook's chronology was the chronology of Aisen's attention, and his attention had moved from Naia to Thane the way a surveyor's eye moves from the broken ground to the structure responsible for the breaking: upward, toward the source. Three days after the healing practicum, a Tuesday, Aisen had been seated in the second row of the ritual theory lecture—a class he attended not for its content, which he could have absorbed from the course text in an evening, but for the proximity it provided to the faculty's operational patterns. Thane had been lecturing on the theoretical basis of binding circles—the geometry of containment, the mathematical relationships between circle radius and Flux density that determined whether a binding held or failed—and the lecture had been competent, organised, delivered with the measured pace of someone who had given the same material enough times to have smoothed it into a performance that required minimal cognitive effort. A good lecture. An adequate lecture. The kind of lecture that a student could attend for an entire term and remember nothing about the person who had delivered it.

Except.

At the forty-minute mark, a knock had interrupted the lecture. A young man in the undecorated grey of an administrative aide opened the door, spoke two sentences that Aisen could not hear from the second row, and left. Thane had excused himself—politely, with the practised ease of a man accustomed to the minor interruptions of institutional life—stepped into the corridor for ninety seconds, and returned. He resumed the lecture at the precise sentence where he had been interrupted. No hesitation, no consultation of notes, no orienting pause. The continuity was seamless.

And wrong.

Because in the ninety seconds of Thane's absence, Aisen had counted—a reflex now, like breathing, the automatic measurement of duration that his years in the tavern and on the street had installed in his nervous system as permanently as the ability to read a room's exits or calculate the weight distribution of a man about to throw a punch. Ninety seconds was nothing. But when Thane returned, his right hand carried the faint scent of seal wax—the particular tallow-and-resin formula used not by the academy's internal correspondence office but by the external courier service that handled communications between the institution and its governing bodies. Aisen knew the difference because Rin had described it to him six weeks ago, a detail extracted from her cataloguing of the academy's logistical infrastructure: internal correspondence used beeswax seals, warm and honeyed; external correspondence used tallow-resin seals, sharp and mineral, with a trace of pine pitch that clung to the fingers and the nose.

Thane had opened an external letter in ninety seconds. Read it. Resealed or disposed of it. Returned. And none of this was remarkable—faculty received external correspondence; it was not prohibited, not unusual, not suspicious in isolation. But the knock had not come during Thane's office hours. It had not come during the intervals between classes when mail was customarily distributed. It had come mid-lecture, from an administrative aide whose grey clothing bore no departmental insignia, which meant either the aide served the general administration or the aide served no department at all.

Aisen had written in the notebook that evening: \emph{Thane. External correspondence. Mid-class delivery. Aide unmarked. Seal wax: tallow-resin, external service. Follow.}