% Auto-generated from 08-Covenant-Scholar.md
% POV: Aisen | Landscape: Mountains | Location: The High Keep

Aisen had seen Soren twice a week for two years. He understood the difference between Soren's normal state and what he saw when the old librarian opened the inner door of the restricted chamber and stepped back to let him through.

The left side of Soren's face had a quality to it that Aisen had no word for immediately — a slight asymmetry in the under-eye area, the skin there appearing both tighter and more translucent than the right, as if the tissue had been threaded with something thin and cold. His movements were unaltered in their precision but slightly slower in their initiation, as though each action required a moment of deliberate will rather than occurring as automatic response. His breathing at rest was very slightly audible.

Aisen filed this and sat down at the table he usually sat at and did not immediately ask.

The restricted chamber was the same as it always was: ceiling-high shelves on three walls, boxes unlabeled, the fourth wall's narrow windows providing flat light. Two candles supplemented the windows, as always. The smell was what it always was in this room: old paper, particular dust, the faint mineral quality of something pre-chemical and dry. He had been in this room approximately forty times. He knew its dimensions and the approximate positions of the boxes he had been permitted to open, and the positions of those he had not.

Soren settled across the table. He placed his hands flat on the surface, a gesture he did not usually make — flat palms down was a stabilization posture, he had observed it in older people when their sense of level was intermittently unreliable.

"Harald says you coordinated the Verin Lane operation cleanly," Soren said. His voice was unaffected. This was apparently where the energy went first: toward maintaining the voice.

"Yes," Aisen said. "Third week of last month."

"He says you used Thrace and kept yourself invisible throughout."

"Yes."

Soren looked at him with the expression that had always, in Aisen's experience, functioned as evaluation rather than communication — taking in information rather than transmitting it. After two years, Aisen had become comfortable with the expression. He waited.

"I asked Harald to send you today because I want to tell you something before I lose the coherence for it," Soren said. "I'm having a good morning, which means I have perhaps an hour before the clarity degrades. I'd rather use the hour well than schedule a better time that may not come."

Aisen kept his hands in his lap. His right hand had done its involuntary motion toward the inner pocket — the notebook — and he left it there. He did not take out the notebook. He listened.

"I have celestial binding," Soren said. The term was a Covenant designation for a class of long-term magical destabilization which the Covenant applied to people whose continued existence was practically convenient but whose continued full capacity was not. Aisen knew what it was from the restricted materials. He had never encountered someone who had it. "It was applied approximately four years ago. I have perhaps three years remaining at functional capacity, after which the degradation becomes rapid."

"Why are you telling me now?" Aisen asked.

"Because there are things that need to be transferred while I can transfer them properly, and I have been delaying under the assumption that I had more time than I appear to have." He moved one hand slightly — a checking motion, feeling the edge of the table. His eyes didn't track the hand. That was a different problem and it was the first one Aisen had not seen before on this visit. "The collection."

He meant the restricted materials. Not all of them — the sanctioned restricted materials, which the Covenant knew held. He meant the second collection: Box J through Box R on the east wall's lower three shelves, which were not logged in the Archives' public catalogue and whose existence the Covenant was unaware of.

Aisen looked at those boxes.

"I maintain irregular access for myself," Soren said. "I am the only person who accesses them. When I am gone — or when I can no longer access them reliably — they will be uncatalogued in a Covenant institution that eventually will be audited. What happens to uncatalogued materials in a Covenant audit is determined by who conducts the audit and what they are told to find."

"They're destroyed," Aisen said.

"Not necessarily immediately. But they are identified, flagged, and at some point classified for removal. There is often a gap of one to three years between identification and physical removal during which the materials still exist and could be relocated." He paused. The slight asymmetry on the left side of his face was more pronounced than it had been five minutes ago. "That gap is what you'd have to work with."

"I don't have access to these materials," Aisen said.

"You have a standing arrangement with me. I can extend your access before I lose the authority to do so. The extension would require a formal notation in the Academy-archive liaison register that creates a legitimate reason for a non-scholar to be in regular contact with the restricted tier." He moved both hands now, resetting them flat on the table with the careful deliberate quality of someone recalibrating. "I would need to designate you as a research correspondent, which is procedurally unusual but not unprecedented. The Covenant would note it. They would decide what it meant based on how much they already knew about you."

"Which is currently nothing," Aisen said.

"Currently nothing," Soren said. "In four years, when you enter the Academy, it will be something. By the time I am gone and the collection needs to be moved, the question of whether your being designated as a research correspondent is a liability or a cover will depend on how well you've managed your own visibility profile in the Academy."

Aisen looked at the boxes marked J through R on the lower east shelves. He could not read the labeling from where he sat — what labeling there was. Some were marked with paper tabs in Soren's handwriting.

"What is in them?" he asked.

"Records of the Resistance Charter's full implementation document," Soren said. "Three complete legal frameworks for post-Covenant governance, from the First, Second, and Fourth attempts. Documentation of the magical suppression program from its beginning, including the methodologies and the names of the original decision-makers. Academic materials on decentralized governance that exist nowhere else in any accessible form. And one object, secured in the tenth box from the left on the middle shelf, that I will describe to you only once and which you should not attempt to access until you have the means to understand what you're looking at."

His voice had gone slightly lower in volume. Not in tone — the precision was intact — but in power. The effort of the long sentences showed slightly in the intervals between them.

"You're telling me this in case you don't get the chance to tell me later," Aisen said.

"I'm telling you this because you're the only person I know who will still be operationally capable in three years time and who understands why these materials matter. Harald understands why they matter but Harald is sixty-one and the celestial binding process is not exclusive." He stopped. His left hand moved off the table briefly, finding the edge of the chair's arm. "If neither of us is capable, these materials become lost. Not destroyed immediately — lost. Which means they might be recovered by someone who can use them as tools for the work, or might be found by someone who brings them to the Covenant's attention. The probability distribution on that outcome is not good."

Aisen looked at the boxes.

He was sixteen. He had conducted three operations, coordinated a counter-messaging campaign, and run observation and information-gathering for the better part of two years. He had a functional working understanding of how the network operated and how to move through the city without becoming visible.

He had not been asked to carry anything before.

The word \emph{carry} arrived in his thinking from somewhere adjacent to conscious decision — he had not used it in any of the sentences he'd spoken — but it was the right word. Soren was asking him to carry this. To hold the knowledge of what existed in Box J through R and to hold the means of accessing it and to hold the responsibility for moving it when the time came. Not to act now. To carry the knowledge forward until acting became possible.

His father had carried something similar, he thought, for most of their shared life.

He had not previously understood what that weight was made of.

"I'll need to understand the liaison notation process," he said. "And the designation you mentioned — research correspondent. I need to know what questions it will raise when I enter the Academy."

Soren's left side of his face had gone slightly slack — barely detectable, a two-percent change in the way the skin held its tension near the cheekbone. The hour was ending.

"That," Soren said, "I'll explain next time. I'm going to need to sit quietly for a few minutes."

Aisen stood. He did not immediately leave. He stood at the table for a moment, with the boxes marked J through R in his peripheral vision and Soren's hands still flat on the table and the candle burning low in its holder, and understood that the meeting had ended in the way that many things ended: without ceremony, in the middle of something rather than at its conclusion, in a room that looked exactly the same as it had when he'd walked in and that was somehow permanently different from the room it had been before.

He let himself out through the bone-token door.

The Archives' main hall was cool and quiet. He walked to the catalogue shelves, selected a volume on pre-Covenant administrative law that he had been meaning to read for several months, found a chair near the window with the best afternoon light, and opened it.

His hands were steady.

His chest, slightly below the sternum, had a sensation that was not quite pain and not quite pressure and was not something he had a name for, which meant he filed it under \emph{physical register} alongside the other involuntary responses he had catalogued but not yet fully decoded.

He read twelve pages before the end of the reading period. He retained approximately six of them.
