% Auto-generated from Chapter-26-Conscription-Separation.md
% Source: C:\Users\PCGAME\Desktop\story&universes\chain-weapon-story\finished-manuscript\manuscriptbaseforchapters\Chapter-26-Conscription-Separation.md

\chapter{Conscription \& Separation}
\renewcommand{\CurrentChapterNum}{26}
\POV{\glyphAisen}{Aisen}

The names were read in the refectory at the seventh morning hour, which was the hour reserved for announcements that the institution considered too important for the day's periphery and too disruptive for the day's centre—the administrative slot that existed between breakfast and first lecture, when the student body was assembled and fed and therefore docile, the animal contentment of a full stomach serving the institution's purpose the way it always served the purposes of those who understood the relationship between compliance and digestion. Aisen knew the hour. He had catalogued the pattern over two years: disciplinary actions at the seventh hour, schedule changes at the seventh hour, the announcement of the winter examinations at the seventh hour. The things that could not be questioned were delivered when the audience was least inclined to question.

The officer was not academy faculty. This was the first detail Aisen registered—before the words, before the content, the uniform wrong for the hall. Covenant military, not institutional staff. A woman in her forties with the close-cropped hair and weathered skin of someone who had spent years outside stonework, whose face carried the particular compression of a person who delivered orders that other people would have to live with. Her uniform was field-grade grey, unadorned except for the Covenant's sigil at the collar—the interlocking rings that denoted jurisdictional authority—and she held a folio in her left hand with the casual grip of someone who had held many folios and felt nothing for any of them.

"By order of the Covenant's Northern Command, the following students are hereby conscripted for border service under Emergency Provision Fourteen, Section Three. Conscripts will report to the eastern courtyard by the ninth hour with personal effects not exceeding one pack. Institutional records will be updated. Academic standing is preserved but suspended for the duration of service."

She opened the folio. The pages inside were organised in the alphabetical architecture of bureaucracy—surnames sorted, first names subordinated, each identity reduced to a line of ink in a sequence that converted individuals into a list and a list into a logistics problem, the transformation from person to unit accomplished not through violence but through the clerical precision of a system that had been converting people into numbers for longer than any of the people in the room had been alive.

The names began. Aisen did not hear the first six because his mind had already divided—one part listening, the other calculating. Emergency Provision Fourteen. He had encountered the provision in the restricted archive, in Rin's copied files: the legal mechanism by which the Covenant could conscript academy students during periods of border escalation, bypassing the standard three-year completion requirement, converting scholars and trainees into soldiers with no more ceremony than a name read from a list. The provision had been invoked eleven times in the last century. Each invocation corresponded to what the archive's dry language called a "period of territorial stress"—the bureaucratic euphemism for war.

"Korv, Aisen."

His name arrived the way all the names arrived: flat, dispassionate, one syllable-set among dozens, the officer's voice maintaining its administrative cadence without emphasis or inflection, each name receiving exactly the same tonal weight as every other name, because the system that produced the list did not distinguish between the names on it. He was not Aisen who had read the restricted archive or Aisen who ran a network or Aisen who carried a chain-spear that no other student at the academy possessed. He was Korv, A., a line of ink, a body that the Covenant required at a location that the Covenant had determined, performing a function that the Covenant would define.

The refectory was silent except for the officer's voice. Two hundred students sat at the long tables with the stunned immobility of people receiving information that reorganised their immediate future—the particular stillness that was not calm but its opposite, the body's response to a threat too abstract for the musculature to address, the mind processing displacement while the hands held spoons and the breakfast cooled in its bowls.

"Northern Garrison Twelve. The Ashenmarch." The officer read the assignment without looking up. "Report to convoy staging."

\emph{The Ashenmarch.} The word entered Aisen's mind and found, already waiting for it, the memory of Harald's voice from years ago—the evening talk at The Crossing, the fire low, his father's hands wrapped around a mug as he described the garrison roads that fed soldiers northward, the supply wagons that passed through the crossroads twice a season carrying grain and iron and letters that were not always answered. \emph{The soldiers who come through here are heading north, most of them,} Harald had said, \emph{and the ones who come back are not the same ones who went.} At twelve, Aisen had understood the sentence as observation. At eighteen, sitting in the refectory with his name on a list, he understood it as prophecy.