% Auto-generated from 09-Final-Warning.md
% POV: Aisen | Landscape: City | Location: The Council Chamber

Harald poured two cups of water and set one in front of Aisen and kept one for himself, and Aisen understood from this gesture alone that the conversation was going to last several hours.

It was nine in the evening. They were in the flat's main room, which was the room they always used for things that mattered, and the small differences from a normal training session were: the hour (late), the water (instead of tea), and the fact that Harald had said nothing since Aisen sat down. The silence had lasted approximately three minutes and had the textured quality of something being organized rather than something being withheld.

The city outside the window was doing the thing it did at nine on a weekday: it had quieted from the market-hour cacophony to a lower register of late movement, cart-wheels over stone and distant voices from windows left open against the warmth. These were sounds Aisen had woken to and slept to for his entire life and they had the unremarkable familiarity of things you stop hearing until something provides contrast.

Harald's hands were flat on the table.

"Two months," Harald said.

"Yes," Aisen said.

"In two months, you'll enter the Academy. Once you do, I'll need to reduce contact significantly. Not eliminate — I'll still be reachable through the network's secondary channels — but I won't be able to call for meetings. I won't be able to run training sessions. You'll be operating on what you have."

"I know," Aisen said.

"I want to tell you the probability estimates before you go," Harald said. "Not to discourage you. Because you should enter with accurate information rather than the kind of selective optimism that makes you under-prepare for what's actually likely."

Aisen's hands were in his lap. He registered, briefly, a faint increase in his heart rate, which he noted without acting on. Harald rarely quantified. When he did, it usually meant the numbers were not good.

"Thirty percent probability of identification and arrest within the first two years," Harald said. "The Covenant's surveillance at the Academy is more sophisticated than in the city networks. They have informants at every level of student life. They run pattern recognition across academic records, association patterns, and behavioral anomalies. If you make a misstep in the early period, you'll be flagged." He paused. "If you're arrested in the Academy, the first consequence is interrogation. The second, if you retain enough utility to be worth converting, is an attempt to turn you. If that fails, execution."

Aisen kept his hands in his lap.

"If you get through the first two years without arrest," Harald continued, "the probability shifts to approximately sixty percent that you'll be arrested and executed before your thirtieth year. The work becomes more visible as you become more capable. Eventually the visibility is incompatible with safety."

"What's the remaining forty percent?" Aisen asked.

Harald looked at him with the specific expression that meant he was deciding how to say something, which was different from the expression that meant he'd decided. "The remaining forty percent includes outcomes ranging from continued operation at reduced capacity, to exile, to outcomes that involve surviving the Academy period in a way I can't currently model because I don't know what the Academy period will produce."

"That's not forty percent," Aisen said.

"No," Harald agreed. "The honest number is that the forty percent is not modeled. I don't know what happens in it."

Aisen looked at the water in his cup. The light caught it slightly — the table lamp was set to the side, at the angle Harald used for evening study sessions, and the geometry of the cup and the water and the lamp angle produced a small crescent of reflection on the table's surface near his hand.

"Do you think we can win?" he asked.

Harald's left hand moved — a very small motion, thumb against table edge, the motion he made when he was considering. "What does winning mean?" he said, and this was not a deflection; this was the question.

"Break the Covenant's control," Aisen said. "Completely. Not disrupt operations. Not create partial resistance. End the system."

Harald was quiet for a moment. The city outside had further quieted. The water cup on the table was perfectly still now.

"Yes," Harald said.

Aisen waited.

"I believe the Covenant is structurally vulnerable in a way its architects didn't understand when they built it," Harald said. "They built it around the assumption that human need for security would perpetually outweigh human desire for self-determination. That's a true assumption in stable conditions. It stops being true when conditions destabilize enough that the security the Covenant provides becomes visibly inadequate. They've built several suppression mechanisms on the assumption that they can always control the conditions, which is different from being capable of controlling the conditions when something they didn't model occurs."

"What's the something they didn't model?" Aisen asked.

"I don't know," Harald said. "That's what I've trained you to find. Not to find me the answer — to build the position where when the unmodeled event happens, there is a distributed network of people capable of recognizing what they're looking at and responding to it faster than the Covenant can recover." His hands moved together now, not flat but folded, the posture he used for things he had thought about for a long time. "That's what the last seven years have been. Not training you to fight the Covenant. Training you to recognize when the moment exists."

Aisen filed this. 

"I want to make sure I understand what you're saying," he said. "You're not saying: go to the Academy, run your operations, build the network, and eventually those things will accumulate into victory. You're saying something happens — something we don't know yet — and the work we've done prepares us to respond to it when it does."

"Yes," Harald said.

"That's a different thing than I thought I was doing," Aisen said.

"Yes," Harald said again. "Most people who do this work think they're doing the first kind. A few of them are right. Usually the ones who get executed believing they're winning are doing that variant. The second kind is slower, harder to maintain motivation for, and more likely to eventually produce the outcome."

Aisen sat with this for a moment. Outside, a cart passed — the slow hollow knock of iron wheel on stone, diminishing south.

"You'll attract people," Harald said. "Followers, in the practical sense. People who believe what you're doing is right and want to be part of it. Some of them will follow you into operational situations and some of them will die because of decisions you make. You should enter the Academy already having accepted that cost."

"I haven't accepted it," Aisen said.

Harald looked at him.

"It's a statement of fact," Aisen said. "I haven't accepted it. I don't think I'm capable of accepting it before it happens. I can recognize intellectually that it will be the case. That's different."

Harald considered this for longer than he usually paused. "That's honest," he said. "Most people claim they've accepted it. That's the wrong posture — it makes them careless with others. The right posture is knowing it will happen and operating with that knowledge present without being paralyzed by it. What you're describing is closer to the right posture than what most people come in with."

They sat for another hour. Harald covered contingency procedures, the Academy's informant structure as he understood it, the two contact points Aisen could use in the city if the network's secondary channels were compromised, and what to do in the specific scenario where he was flagged but not yet arrested — the window of behavior that could either resolve the flag or confirm it, and the difference between them.

When Harald finished speaking, he picked up his water cup and drank the rest of the water. Aisen did the same.

"When will I see you again?" Aisen asked.

"In the city, probably never," Harald said. "I'll be through the secondary channel when you need me. You'll know the signal." He set his cup down. His face had the quality of something that had finished rather than something that had stopped. "The rest is yours."

Aisen stood.

He stood for a moment at the table in the room where he had been trained for seven years, which he was certain he would not see again in the same configuration after tonight — not because it would change, but because the next time he came here, if he came here, the relationship the room represented would have already ended.

He said: "I'll do the work."

Harald said: "I know."

Aisen let himself out.

The city at ten-thirty was cool and quiet and completely indifferent to the conversation that had just concluded, which was appropriate. He walked home through streets he knew from seventeen years of living in them, and the walking was the same as it always was and the streets were the same as they always were and only he was different.

He was not sure, precisely, how he was different. That was something he would understand in retrospect, when he had enough distance to see the shape of the change.

He went to bed and lay still for forty minutes and then slept, because he had learned to sleep on demand even when his chest was doing the unnamed-sensation thing, and the discipline of sleeping on demand was one of the skills he understood, tonight, he was going to need.
