% Auto-generated from Chapter-13-Academy-Arrival.md
% Source: C:\Users\PCGAME\Desktop\story&universes\chain-weapon-story\finished-manuscript\manuscriptbaseforchapters\Chapter-13-Academy-Arrival.md

\chapter{Academy Arrival}
\renewcommand{\CurrentChapterNum}{13}
\POV{\glyphAisen}{Aisen}

The gate was iron and old and taller than any doorway Aisen had seen—two panels of scrollwork bolted to stone pillars that rose into the morning haze like the pronouncements of an argument too large to have been settled in any single generation. The ironwork depicted scenes he could not fully parse from below: scholars at lecterns, soldiers in formation, a figure that might have been a god or a dean standing above both with hands spread in a gesture that could have meant benediction or ownership, depending on the century of the viewer. Rust had colonized the lower hinges and crept upward in orange tributaries that followed the seams between panels, and the whole structure smelled of mineral decay and lamp oil and the particular institutional authority of a place that had outlasted enough people to stop noticing them as individuals.

Aisen stood before it with a single pack on his shoulders and the sensation of having arrived at the edge of a vocabulary he did not yet speak.

Harald was beside him. His father had said almost nothing during the final day of travel—the wagon ride from the border relay, the transfer to the academy's supply road, the last stretch on foot through the campus grounds where the buildings thickened and rose around them like the walls of a narrowing corridor. Harald's silence was not the absence of speech but the presence of too many things to say, compressed into a stillness that Aisen recognized because he had inherited it: the Korv men's particular talent for feeling deeply and expressing sparingly, as though emotion were a resource too valuable to spend without certainty it would be received at full weight.

They stood together before the gate and said nothing, and the nothing they said was enormous.

"Your room assignment will be in the eastern dormitory," Harald said, finally. He was reading the notice board beside the gate, a slab of dark wood covered in pinned papers that fluttered in the thin breeze like captured birds. His voice was practical, directed at logistics, because logistics were the ground on which Harald's affection had always built its structures. "Third floor. Room seventeen."

"I see it."

"You'll want to claim the bed nearest the window. Better air. And you can see who approaches the building from the courtyard."

Aisen almost smiled. Three years at The Crossing's crossroads, reading arrivals through glass—his father could not stop teaching him to watch doorways even now, even here, even at the threshold of a place designed to teach him things Harald had never learned and never would.

"I'll take the window bed," Aisen said.

Harald nodded. He adjusted the collar of his coat—a gesture that served no practical purpose, the collar having been perfectly positioned already, the motion a translation of something internal into something physical because the internal thing could not be spoken in full. Aisen waited. He had learned, over seventeen years of studying the most important text in his life, that his father's significant statements arrived after physical preamble, the body clearing its throat before the heart could speak.

"You have Tarovin's letter," Harald said.

"In the pocket." Aisen touched his chest where the folded paper rested against his ribs, Tarovin's reluctant blessing committed to ink in a scholar's precise hand—a letter of recommendation that was simultaneously genuine and strategic, every word both true and calculated, because Tarovin understood that in institutions the truth was never sufficient and had to be presented in the institution's own dialect to be heard.

"And you have your notebook."

"Always."

"Then you have what you need." Harald turned to face him, and the turning was deliberate, a reorientation of the body that signaled a shift in register from practical to essential. His eyes—gray, steady, the eyes of a man who had spent thirty years reading rooms and had learned that the most important readings were the ones closest to home—held Aisen's with a directness that was itself a form of tenderness. "Listen more than you talk. Watch more than you act. And remember that the people who explain the rules first are rarely the ones who made them."

Aisen felt the weight of the statement settle into his chest alongside Tarovin's letter, another piece of folded wisdom carried close to the heart. He wanted to say something that matched it—something that acknowledged the years of lessons delivered through bar conversation and quiet demonstration, the education that had no curriculum but had been more rigorous than anything the academy's iron gate could promise. The words assembled in his throat and stopped there, too large for the passage they would need to travel, and what came out instead was simpler and truer: "Thank you. For everything."

Harald's expression shifted by degrees that a stranger would have missed entirely—a softening around the eyes, a fractional release of tension in the jaw, the micro-movements of a man allowing himself one unguarded moment before reassembling the architecture of restraint that had carried him through a life of difficult observations. He placed his hand on Aisen's shoulder. The grip was firm, and it lasted exactly long enough to say what neither of them would say aloud, and then it released, and the moment was over, and the gate was waiting.

"Go," Harald said. "Before I start giving you advice about the food."

Aisen went.