% Auto-generated from 07-Network-Test.md
% POV: Aisen | Landscape: City | Location: The Fortress Hall

At a quarter past six in the morning, Aisen was sitting on a stone ledge on the east side of Verin Lane, eating a piece of bread that had gone slightly stale overnight and watching a door.

The door belonged to a workshop owned by a man named Cael, who was not part of the resistance network but whose apprentice was. The apprentice's name was Prend. He was sixteen. He had been on the recruitment office's pre-identification list for fourteen days, which was how long these things took from flag to contact. This morning, at approximately eight o'clock, two Covenant recruitment officers were scheduled to arrive at his door.

Aisen had forty-one minutes.

The problem was: he did not know if the information had reached Prend. He had routed it through Thrace, the merchant, who ran a second-hand trade network that moved through every district in the lower city three days a week. The point of using Thrace was that Thrace's route was invisible to the Covenant — a fabric merchant delivering end-of-bolt scraps to residential households was not a courier system worth surveilling. But Thrace's last confirmed pass through this block was two days ago. Aisen had no indication whether Thrace had made an additional pass, whether the information had been received, or whether Prend had had time to read and act on it.

The bread was cold. The stone was cold. His jacket had one torn pocket seam that let air in at the side of the left hip, which was a problem he had been meaning to fix for several weeks.

He watched the door.

Forty-one minutes was not enough time to approach the door directly, because if the Covenant had moved up their arrival schedule — which happened with approximately one-in-six probability, based on the pattern Corvus had confirmed over six previous recruitment drives — then Aisen approaching the door would create an observable association between him and Prend that did not currently exist. The visible chain had to stay clean.

He checked the bread and decided it was not worth finishing.

The lane's morning deliveries had started at six: two women with market baskets had come up from the south side, a cartman had moved a load of empty crates past going north, and a man Aisen didn't recognize had walked through without stopping at any door. The morning traffic pattern was normal. Normal traffic pattern meant the Covenant team hadn't arrived early and hadn't posted a watch overnight, both of which were reassuring.

What was not reassuring was the candle in Prend's second-floor window.

Prend's household woke at approximately seven on workdays — the workshop opened at half-past seven, and there was a predictable pattern of light in the upstairs window that tracked the household's routine. At a quarter past six, the candle in the second-floor window should not be lit.

Either Prend was awake early — possible, if he'd received the information and was processing it, which would be a reasonable response to being told that enlistment into the Covenant military had an average field-deployment lifespan of fourteen months — or someone else in the household was awake early for reasons unconnected to Prend or the recruitment drive.

Or the Covenant team had already been there.

Aisen's right hand had done a thing where it moved slightly toward his jacket's inner pocket — where the notebook was — before he had consciously decided to take notes. He recognized this as an agitated response and did not take out the notebook. He kept his hands in his lap and kept the door in his sightline and counted backward from forty in intervals of seven, which was a practice that had the effect of settling the impulse to act before he had enough information to act usefully.

Forty. Thirty-three. Twenty-six. Nineteen. Twelve. Five. He started again.

The door opened.

A woman in her fifties — the workshop owner's wife, whose name he didn't know but whose face he recognized from three weeks of observation — came out and placed an empty ceramic bowl on the step. This was a household signal: it meant breakfast was ready, and someone was being called. She looked up and down the lane without apparent purpose or alarm, and went back inside.

Normal.

The candle in the second-floor window moved.

Someone had picked it up. Then it was set down in a different position — left side of the sill rather than center. Aisen watched this and ran his thumb along the bottom edge of the stone ledge until he found a rough patch that he could press his nail against, because the physical sensation of pressure helped him hold still and not visibly react.

Left side of the sill.

He did not know if Prend had been briefed on a candle code. He had not briefed Prend on anything directly — the information had gone through Thrace and through a second-layer courier who had received instructions from Lyssa. He did not know if Lyssa had included signal instructions or if the candle position was coincidence.

He was going to have to operate without knowing.

That, Harald had told him, was most of the operational condition. You made decisions on the information available and you did not demand certainty before acting, because the demand for certainty was itself a form of inaction and inaction had costs that were as real as the costs of wrong action. The question was not \emph{am I sure} but \emph{am I sure enough.}

He ate the rest of the bread.

At six forty-seven, Thrace came around the corner from the south side of the lane. He was carrying a cloth bundle under one arm — end-of-bolt scraps, the faded blue-and-grey pattern that Aisen had seen him deliver to three different households in the past two weeks. He stopped at the workshop door. Knocked twice. Waited. The woman answered, accepted the cloth bundle, handed him a small coin. He continued north up the lane at the same pace he always maintained.

As he passed Aisen's ledge he said, very quietly, without turning his head or altering his pace in any detectable way: "Yesterday morning. He's read it twice."

Aisen kept his face precisely neutral.

Thrace continued up the lane and turned left, out of sight.

He had thirty-one minutes.

He stood from the ledge and walked south, in the direction Thrace had come from, at a pace calibrated to be a person with a morning errand and a clear destination. His feet found the familiar cadence of it — a particular rhythm he had practiced until it was automatic, the gait of someone who lived in this district and was not thinking about what was behind them.

He turned right at the first intersection and took the longer way back.

The information had reached Prend. Prend had read it twice, which was sufficient engagement to indicate he'd taken it seriously rather than dismissed it. He now knew what the recruitment drive would mean in practice — the deployment statistics, the casualty rates, the contractual terms that were not said during the enrollment interview. He knew that refusal had procedural consequences which were theoretically severe and practically modest if navigated correctly. He knew that the network existed, in general terms, though not how to access it, which was deliberate: knowledge of existence without knowledge of access was protection, not a liability.

When the Covenant officers arrived at eight, they would find a young man who had answers to every question they would try to use as leverage.

Aisen did not know what Prend would decide.

That was correct. The purpose of the operation was not to make Prend's decision for him. It was to ensure he had the information that the Covenant relied on him \emph{not} having. What Prend chose to do with accurate information was Prend's right, not Aisen's.

He'd had this thought before — it had appeared first when Harald had said the network's purpose as widening the range of what people could choose, rather than dictating what they should choose. At fourteen, Aisen had found this philosophically satisfying and practically somewhat abstract. At fifteen and a half, having spent three weeks watching Verin Lane and tracking the recruitment drive's mechanics, he found it still philosophically satisfying and marginally less abstract.

The lane behind him was quiet.

He walked the long way home through the city waking up around him — carters and early market workers and the smell of bread from the bakery on Merchant's Cross and the damp cold of the river two streets west, which the wind was bringing in from the east — and by the time he reached the door to the flat where he slept, the sky had gone from grey to the particular pale yellow that meant the sun was up but not yet high enough to be warm.

He went upstairs and wrote three lines in the notebook. Not a full record — just the operational notation that would be referenced against Harald's debriefing later. Then he lay down for the hour he had before he was supposed to be at the Archives, and stared at the ceiling's water-stain pattern, which was a fixed geographic feature of his sleep space that he had been using as an anchor for ordered thinking since he was eleven.

The operation had worked. The operation had worked because Thrace had done his job and Lyssa had done hers and Corvus had provided the right logistical detail at the right time. Aisen had coordinated and observed and had not been seen and had not done anything that placed a visible thread between himself and any part of the counter-messaging chain.

This was the correct outcome.

He closed his eyes.

\emph{From a position, you can operate.} His father's chess sentence. He had a better position now than he'd had three weeks ago — not a resolved position, but a better one. He had run a small thing, and the small thing had worked, and working meant he understood more about how the network functioned under operational conditions than he had understood before.

He had thirty-eight minutes before the Archives.

He let himself sleep for thirty-five of them.
