\documentclass[11pt,openany]{book}
\usepackage[margin=0.95in]{geometry}
\usepackage{fontspec}
% Prefer TeX Gyre fonts if available; otherwise fall back to standard LaTeX fonts
\usepackage{microtype}
\usepackage[table]{xcolor}
\usepackage{hyperref}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage{array}
\usepackage{enumitem}
\usepackage{titlesec}
\usepackage{fancyhdr}

% Set fonts when available, else use safe LaTeX package fallbacks
\IfFontExistsTF{TeX Gyre Pagella}{%
  \setmainfont{TeX Gyre Pagella}%
}{%
  \usepackage{mathpazo}% Palatino-like fallback
}
\IfFontExistsTF{TeX Gyre Heros}{%
  \setsansfont{TeX Gyre Heros}%
}{%
  \usepackage{helvet}% Helvetica fallback for sans-serif
  \renewcommand{\sfdefault}{phv}% set helvet as sans default
}

\definecolor{Ink}{HTML}{1F2933}
\definecolor{DeepSea}{HTML}{0B3C5D}
\definecolor{Accent}{HTML}{D97706}
\definecolor{Soft}{HTML}{F5F8FA}

\color{Ink}

\hypersetup{
  colorlinks=true,
  linkcolor=DeepSea,
  urlcolor=Accent,
  pdftitle={Tseluca Offshore Playbook},
  pdfauthor={Petroleum Engineering Market Prospects}
}

\setlist[itemize]{itemsep=0.35em, topsep=0.35em, leftmargin=1.6em}
\setlist[enumerate]{itemsep=0.35em, topsep=0.35em, leftmargin=1.8em}

\titleformat{\chapter}[display]
  {\normalfont\sffamily\bfseries\color{DeepSea}}
  {\filleft\fontsize{42}{42}\selectfont\color{Accent}\thechapter}
  {0.8ex}
  {\titlerule[0.9pt]\vspace{0.8ex}\filright\Huge}
  [\vspace{0.6ex}\titlerule]

\titlespacing*{\chapter}{0pt}{-6pt}{22pt}

\titleformat{\section}
  {\normalfont\sffamily\bfseries\Large\color{DeepSea}}
  {\thesection}
  {0.7em}
  {}

\titleformat{\subsection}
  {\normalfont\sffamily\bfseries\color{DeepSea}}
  {\thesubsection}
  {0.6em}
  {}

\pagestyle{fancy}
\fancyhf{}
\fancyhead[LE,RO]{\sffamily\small\color{DeepSea}Tseluca Offshore Playbook}
\fancyhead[LO,RE]{\sffamily\small\color{Ink}\leftmark}
\fancyfoot[C]{\sffamily\small\thepage}
\setlength{\headheight}{14pt}
\raggedbottom

\title{Tseluca Offshore Playbook\\\large What To Do, Where To Do It, and How To Get Hired Fast}
\author{Built for ISPTEC Student Execution}
\date{April 2026}

\begin{document}

\frontmatter
\begin{titlepage}
\thispagestyle{empty}
\colorbox{Soft}{
  \parbox{\dimexpr\textwidth-2\fboxsep\relax}{
    \vspace{1.1cm}
    {\sffamily\bfseries\LARGE\color{DeepSea}TSELUCA OFFSHORE PLAYBOOK\par}
    \vspace{0.35cm}
    {\sffamily\large\color{Ink}What To Do, Where To Do It, and How To Get Hired Fast\par}
    \vspace{0.75cm}
    {\sffamily\color{Accent}\rule{0.32\textwidth}{1.4pt}\par}
    \vspace{0.85cm}
    {\large\bfseries Built for ISPTEC Student Execution\par}
    \vspace{0.3cm}
    {April 2026\par}
    \vspace{1.0cm}
    \textbf{Purpose}\par
    A practical execution book for offshore FIFO employability,\\
    petroleum-focused technical growth, and salary progression\\
    in Angola and regional markets.
    \vspace{1.2cm}
  }
}
\vfill
{\small\color{Ink}Repository: petroleum-engineering-market-prospects}
\end{titlepage}

\chapter*{How To Use This Book}
This book is an execution manual for one goal: get into a petroleum-focused offshore rotational role quickly, with a strong salary path and low unemployment risk after graduation.

Use this method:
\begin{enumerate}
  \item Read Chapters 1 to 3 in one sitting.
  \item Build your personal checklist from Chapters 4 to 7.
  \item Execute weekly using Chapter 8 scorecards.
  \item Review and update your plan every Sunday.
\end{enumerate}

\tableofcontents

\mainmatter

\chapter{Your Goal and Best Strategy}

\section{Your stated target}
\begin{itemize}
  \item Work offshore in FIFO rotation lifestyle.
  \item Earn good salary with upward trajectory.
  \item Avoid long unemployment after graduation.
  \item Stay in petroleum-engineering-focused roles.
\end{itemize}

\section{Best combined route from repository guidance}
\begin{enumerate}
  \item Pick one primary technical path: Drilling or Production/FPSO.
  \item Complete mobilization certifications early.
  \item Apply in waves to multiple companies in parallel.
  \item Enter via high-employability routes first (service companies, then operators/FPSO).
  \item Build technical depth and move up by proven performance.
\end{enumerate}

\section{Career architecture}
\begin{itemize}
  \item Stage 0: final year and graduation launch.
  \item Stage 1: first role (0 to 6 months after graduation).
  \item Stage 2: specialization growth (Years 1 to 2).
  \item Stage 3: salary acceleration (Years 2 to 5).
  \item Stage 4: premium offshore technical roles (Year 5+).
\end{itemize}

\chapter{What To Do First: 90-Day Launch Plan}

\section{Days 1 to 15}
\begin{itemize}
  \item Choose primary path: Drilling or Production/FPSO.
  \item Choose backup path: HSE-support or Completions-support.
  \item Build target company sheet (minimum 20 companies).
  \item Open certification savings plan and fund it monthly.
  \item Join SPE student membership.
\end{itemize}

\section{Days 16 to 45}
\begin{itemize}
  \item Complete offshore medical.
  \item Book OPITO BOSIET with HUET.
  \item Start IWCF Level 1 preparation.
  \item Build CV and LinkedIn profile (Portuguese and English ready).
\end{itemize}

\section{Days 46 to 75}
\begin{itemize}
  \item Complete BOSIET/HUET.
  \item Complete H2S awareness (if required by targets).
  \item Complete IWCF Level 1.
  \item Build first technical mini-project (Excel or Python).
\end{itemize}

\section{Days 76 to 90}
\begin{itemize}
  \item Submit first application wave (10 to 15 strong applications).
  \item Run mock interviews weekly.
  \item Attend one SPE/ANPG/company networking event.
  \item Start NEBOSH IGC if budget and role strategy support it.
\end{itemize}

\chapter{Certifications: What, Where, and How}

\section{Priority order}
\begin{enumerate}
  \item Offshore medical certificate.
  \item OPITO BOSIET with HUET.
  \item IWCF Level 1.
  \item H2S Awareness (based on project and employer).
  \item NEBOSH IGC (strong multiplier, especially for HSE-linked roles).
  \item IWCF Level 2 (often best with company sponsorship after hiring).
\end{enumerate}

\section{Angola and regional Africa location map}
\begingroup
\footnotesize
\setlength{\tabcolsep}{4pt}
\rowcolors{2}{Soft}{white}
\begin{longtable}{>{\raggedright\arraybackslash}p{0.19\textwidth} >{\raggedright\arraybackslash}p{0.31\textwidth} >{\raggedright\arraybackslash}p{0.25\textwidth} >{\raggedright\arraybackslash}p{0.19\textwidth}}
\toprule
\rowcolor{DeepSea!14}
\textbf{Certification} & \textbf{Angola (Primary)} & \textbf{Africa Backup} & \textbf{Verify On} \\
\midrule
Offshore Medical & Clinica Multiperfil, Hospital Sagrada Esperanca, International SOS Angola & South Africa occupational health providers; Nigeria offshore-medical providers & Employer approved list \\
OPITO BOSIET/HUET & PETROTEC (Luanda), ICAT (Luanda; verify current status) & Nigeria (Lagos), South Africa (Cape Town/Johannesburg), Namibia (limited market) & OPITO approved listings \\
H2S Awareness & PETROTEC or employer induction programs & Nigeria/South Africa safety providers & Employer acceptance \\
IWCF Level 1/2 & Angola centers in IWCF pathways (live check needed) & Nigeria and South Africa IWCF centers & IWCF center directory \\
NEBOSH IGC & Usually online from accredited partners while based in Angola & South Africa/Nigeria providers & NEBOSH learning partners \\
IOSH Managing Safely & Online accredited providers & South Africa/Nigeria providers & IOSH provider directory \\
\bottomrule
\end{longtable}
\endgroup

\section{Direct certification links (clickable)}
Use these official directories and providers to find current approved centers and register safely.

\begin{itemize}
  \item OPITO approved training providers: \url{https://www.opito.com/find-a-training-provider}
  \item IWCF accredited training and exam centers: \url{https://www.iwcf.org/find-centres/}
  \item NEBOSH learning partners: \url{https://www.nebosh.org.uk/study-with-us/find-learning-partner/}
  \item IOSH training provider network: \url{https://iosh.com/training/}
  \item SPE membership and resources: \url{https://www.spe.org/en/membership/}
  \item Astutis (NEBOSH/IOSH online courses): \url{https://www.astutis.com/}
  \item British Safety Council courses: \url{https://www.britsafe.org/training-and-learning/}
\end{itemize}

Angola and regional institutions to monitor:
\begin{itemize}
  \item ANPG (events and industry ecosystem): \url{https://anpg.co.ao/}
  \item Sonangol careers and announcements: \url{https://www.sonangol.co.ao/}
\end{itemize}

\section{Anti-scam quality checks before payment}
\begin{itemize}
  \item Verify accreditation on the official awarding body website.
  \item Confirm exact certificate title and validity in writing.
  \item Confirm target employer acceptance before payment.
  \item Confirm all fees (tuition, exam, practical, re-sit, travel).
  \item Review cancellation and refund policy.
\end{itemize}

\chapter{Skills: What To Develop and How To Develop It}

\section{Technical skill stack}
\begin{itemize}
  \item Well control logic and barriers.
  \item Drilling fundamentals: fluids, ECD, casing and cementing basics.
  \item FPSO process and production troubleshooting.
  \item Formation evaluation basics and production data interpretation.
\end{itemize}

\section{Execution method}
\begin{enumerate}
  \item Study 6 to 8 hours per week.
  \item Solve 2 technical problem sets per week.
  \item Build one concept sheet per week from memory.
  \item Explain one technical topic each week in English.
  \item Build mini-project artifacts that prove competence.
\end{enumerate}

\section{Mini-project ideas}
\begin{itemize}
  \item Drilling dashboard: mud weight, ECD, pressure margins.
  \item Production dashboard: rate tracking, decline trend, anomaly notes.
  \item Interview brief: one page case with diagnosis and recommendation.
\end{itemize}

\section{Where to learn abilities (free and paid links)}
\subsection{Free and low-cost technical learning}
\begin{itemize}
  \item SPE webinars and technical content: \url{https://www.spe.org/en/training/}
  \item OnePetro paper library (abstracts and papers): \url{https://onepetro.org/}
  \item NPTEL engineering courses: \url{https://nptel.ac.in/}
  \item MIT OpenCourseWare engineering: \url{https://ocw.mit.edu/}
  \item Schlumberger Oilfield Glossary: \url{https://glossary.slb.com/}
  \item Halliburton technical resources: \url{https://www.halliburton.com/}
\end{itemize}

\subsection{Paid structured courses}
\begin{itemize}
  \item PetroSkills (drilling, production, reservoir, completions): \url{https://www.petroskills.com/}
  \item OETC (industry training catalog): \url{https://www.oetc.com/}
  \item Coursera energy and data courses: \url{https://www.coursera.org/}
  \item Udemy technical tools (Excel/Python): \url{https://www.udemy.com/}
\end{itemize}

\chapter{Job Targeting for Fast Employability}

\section{Company sequence for your goal}
\begin{enumerate}
  \item Service companies first for entry speed and structured training.
  \item Operators and FPSO owners in parallel for higher long-term ceiling.
  \item Use referrals and networking while applying broadly.
\end{enumerate}

\section{Application wave model}
\begin{itemize}
  \item Wave size: 10 to 15 applications.
  \item Timing: submit within 2 weeks.
  \item Follow-up: 7 to 10 days after application.
  \item Iterate CV and interview answers after each wave.
\end{itemize}

\section{Direct application links}
\begin{itemize}
  \item TotalEnergies careers: \url{https://careers.totalenergies.com/}
  \item Chevron careers: \url{https://careers.chevron.com/}
  \item Eni careers: \url{https://www.eni.com/en-IT/careers.html}
  \item SLB careers: \url{https://careers.slb.com/}
  \item Halliburton careers: \url{https://jobs.halliburton.com/}
  \item Baker Hughes careers: \url{https://careers.bakerhughes.com/}
  \item TechnipFMC careers: \url{https://careers.technipfmc.com/}
  \item Subsea7 careers: \url{https://careers.subsea7.com/}
  \item Oceaneering careers: \url{https://careers.oceaneering.com/}
  \item Weatherford careers: \url{https://jobs.weatherford.com/}
  \item SBM Offshore careers: \url{https://www.sbmoffshore.com/careers/}
  \item Yinson careers: \url{https://www.yinson.com/careers/}
  \item Azule Energy careers: \url{https://www.azule-energy.com/careers}
\end{itemize}

\section{Documents to keep ready}
\begin{itemize}
  \item Passport or national ID.
  \item Degree/transcript or enrollment proof.
  \item Offshore medical certificate.
  \item BOSIET/HUET certificate.
  \item H2S and IWCF certificates.
  \item CV in PDF and LinkedIn link.
  \item Reference list (2 to 3 contacts).
\end{itemize}

\chapter{Interview and Offer Strategy}

\section{Technical interview preparation}
Focus areas:
\begin{itemize}
  \item Well control basics and pressure logic.
  \item Drilling fluid fundamentals and ECD reasoning.
  \item Production basics: IPR/VLP concepts and troubleshooting thinking.
  \item Offshore safety behavior and incident prevention mindset.
\end{itemize}

\section{Behavioral interview preparation}
\begin{itemize}
  \item Teamwork under fatigue and shift pressure.
  \item Communication in multicultural offshore crews.
  \item Safety-first decisions when production pressure exists.
  \item Learning speed and humility in field operations.
\end{itemize}

\section{Offer evaluation checklist}
\begin{itemize}
  \item Rotation pattern and on/off days.
  \item Base pay + offshore allowances + total package.
  \item Certification sponsorship and training roadmap.
  \item Scope of technical work and growth path.
  \item Team quality and mentorship access.
\end{itemize}

\chapter{Weekly Operating System}

\section{Weekly scorecard}
Score each item from 0 to 1.
\begin{itemize}
  \item 6 to 8 hours technical study completed.
  \item 2 problem sets completed.
  \item 3 networking actions completed.
  \item 5 quality applications or 2 referral actions completed.
  \item 1 interview practice completed.
  \item 1 certification savings transfer completed.
\end{itemize}

Interpretation:
\begin{itemize}
  \item 5/6 = strong week.
  \item 6/6 = elite week.
\end{itemize}

\section{Monthly review}
\begin{enumerate}
  \item Which actions produced interviews?
  \item Which gaps blocked progress?
  \item Which certificate or skill must move up in priority?
  \item What is the next 30-day objective?
\end{enumerate}

\chapter{12-Month Roadmap}

\section{Months 1 to 3}
\begin{itemize}
  \item Finalize path and budget.
  \item Complete medical and book BOSIET/HUET.
  \item Launch networking and profile positioning.
\end{itemize}

\section{Months 4 to 6}
\begin{itemize}
  \item Complete BOSIET/HUET and IWCF L1.
  \item Complete first portfolio artifact.
  \item Run first major application waves.
\end{itemize}

\section{Months 7 to 9}
\begin{itemize}
  \item Tighten interview performance.
  \item Expand applications and referral channels.
  \item Add second portfolio artifact.
\end{itemize}
sz
\section{Months 10 to 12}
\begin{itemize}
  \item Convert interviews to offers.
  \item Choose role with best growth and training profile.
  \item Prepare for mobilization and offshore transition.
\end{itemize}

\backmatter

\chapter*{Closing Note}
The strongest pattern across this repository is simple: people who execute consistently win.

You do not need a perfect plan. You need a clear path, valid certifications, technical proof of competence, and weekly discipline.

Start now. Keep score weekly. Iterate monthly. Apply in waves. Build relationships through real performance.

That is how you get offshore fast and grow into high-compensation petroleum roles.
 
% ------------------------- Relat\'orio Consolidado Inserido -------------------------
\chapter*{Relat\'orio Consolidado: Planeamento de Carreira Offshore}
\addcontentsline{toc}{chapter}{Relat\'orio Consolidado: Planeamento de Carreira Offshore}

\section*{Resumo executivo}
Este relat\'orio consolida o plano de carreira offshore para entrada em Produ\c{c}\~ao (FPSO) ou Perfura\c{c}\~ao (Drilling) desde Maio/2026 at\'e Julho/2028, incluindo proje\c{c}\~oes financeiras (em kwanzas AOA), cronograma mensal de certifica\c{c}\~oes, sugest\~oes de fornecedores em Angola, lista de cargos-alvo, uma cole\c{c}\~ao de mini-projetos pr\'aticos para o seu portf\'olio e passos para migra\c{c}\~ao internacional.

\subsection*{Principais conclus\~oes}
\begin{itemize}
  \item Or\'camento acumulado dispon\'ivel at\'e Jul/2028: 3.200.000 AOA.
  \item Prioridade t\'ipica: Exame M\'edico Offshore, OPITO BOSIET/HUET, IWCF L1, H2S, seguido de um curso core (IWCF L2 para perfura\c{c}\~ao ou OPITO Processing para produ\c{c}\~ao).
  \item Recomenda\c{c}\~ao estrat\'egica: Come\c{c}ar por Produ\c{c}\~ao para maior estabilidade; manter a op\c{c}\~ao de migrar para Perfura\c{c}\~ao ap\'os 3--5 anos se desejar maior remunera\c{c}\~ao.
\end{itemize}

\section{Or\'camento e proje\c{c}\~ao financeira (AOA)}
\subsection{Rendimento e soma acumulada}
\begin{itemize}
  \item Maio/2026 (m\'es inicial): 210.000 AOA.
  \item Junho/2026 a Julho/2028 (26 meses): 115.000 AOA por m\^es.
  \item Total acumulado at\'e Julho/2028: 3.200.000 AOA.
\end{itemize}

\subsection{Desagregamento m\'es a m\'es}
\begin{itemize}
  \item Maio/2026: saldo 210.000 AOA.
  \item Junho/2026: saldo 325.000 AOA.
  \item Julho/2026: realizar Exame M\'edico Offshore (105.000 AOA) -- saldo 335.000 AOA.
  \item Agosto/2026: IWCF Level 1 (185.000 AOA) -- saldo 265.000 AOA.
  \item Set/2026 -- Mar/2027: poupan\c{c}a para BOSIET.
  \item Abr/2027: OPITO BOSIET/HUET (1.110.000 AOA) -- saldo ap\'os compra ~75.000 AOA.
  \item Mai/2027: H2S Awareness (85.000 AOA) -- saldo ap\'os compra ~105.000 AOA.
  \item Jun/2027 -- Jan/2028: poupar para curso core (IWCF L2 ou OPITO Processing).
  \item Fev/2028: curso core (aprox. 1.200.000--1.300.000 AOA) -- pequeno ajuste de fluxo previsto.
  \item Jul/2028: saldo final estimado ~570.000 AOA (reserva para imprevistos).
\end{itemize}

\section{Escolha estrat\'egica: Produ\c{c}\~ao vs Perfura\c{c}\~ao}
\subsection{Resumo}
Produ\c{c}\~ao (FPSO): maior estabilidade e maior n\'umero de vagas em Angola; ideal para come\c{c}ar.

Perfura\c{c}\~ao (Drilling): maior potencial de remunera\c{c}\~ao a longo prazo, mas trabalho por campanhas e maior volatilidade.

\subsection{Recomenda\c{c}\~ao}
Come\c{c}ar por Produ\c{c}\~ao para ganhar experi\^encia e estabilidade; manter as certifica\c{c}\~oes transversais e considerar migra\c{c}\~ao para Perfura\c{c}\~ao ap\'os 3--5 anos se desejar maximizar ganhos.

\section{Lista completa de certifica\c{c}\~oes (prioridade e estimativas)}
\subsection{Base comum (obrigat\'oria)}
\begin{itemize}
  \item Offshore Medical Certificate --- 105.000 AOA (validade: 2 anos).
  \item OPITO BOSIET with HUET --- 1.110.000 AOA (validade: 4 anos).
  \item H2S Awareness --- 85.000 AOA (validade: 1--2 anos).
  \item IWCF Level 1 --- 185.000 AOA (investimento de base, sem expira\c{c}\~ao pr\'atica de conte\u{u}do).
\end{itemize}

\subsection{Cursos core por caminho}
\begin{itemize}
  \item Produ\c{c}\~ao: OPITO Oil \& Gas Processing Operations --- ~1.300.000 AOA.
  \item Perfura\c{c}\~ao: IWCF Level 2 --- ~1.200.000 AOA.
\end{itemize}

\subsection{Multiplicadores (ap\'os contrata\c{c}\~ao)}
\begin{itemize}
  \item NEBOSH IGC --- ~1.150.000 AOA (forte diferencial para fun\c{c}\~oes HSE e supervis\~ao).
  \item IWCF Levels 3 e 4 --- custos elevados e normalmente patrocinados pela empresa.
\end{itemize}

\section{Cronograma t\'atico (m\'es a m\'es)}
\begin{enumerate}
  \item Maio/2026: abrir conta de poupan\c{c}a para certifica\c{c}\~oes; juntar 210.000 AOA.
  \item Jun/2026: consolidar poupan\c{c}as (saldo 325.000 AOA).
  \item Jul/2026: Exame M\'edico Offshore (105.000 AOA).
  \item Ago/2026: IWCF Level 1 (185.000 AOA).
  \item Set/2026--Mar/2027: poupan\c{c}a para BOSIET.
  \item Abr/2027: OPITO BOSIET/HUET (1.110.000 AOA).
  \item Mai/2027: H2S Awareness (85.000 AOA).
  \item Jun/2027--Jan/2028: poupan\c{c}a para curso core.
  \item Fev/2028: curso core (IWCF L2 ou OPITO Processing).
  \item Mar--Jul/2028: candidaturas em ondas e prepara\c{c}\~ao para entrevistas.
\end{enumerate}

\section{Fornecedores e verifica\c{c}\~oes (Angola e links oficiais)}
\begin{itemize}
  \item OPITO: verifique centros aprovados em \url{https://www.opito.com/find-a-training-provider}.
  \item IWCF: lista de centros em \url{https://www.iwcf.org/find-centres/}.
  \item NEBOSH: learning partners em \url{https://www.nebosh.org.uk/study-with-us/find-learning-partner/}.
  \item Sugest\~oes locais: OMITC Academy, PETROTEC, CFWC (INP), International SOS, Cl\'inica Multiperfil.
\end{itemize}

\section{Posi\c{c}\~oes-alvo e canais de recrutamento}
\subsection{Produ\c{c}\~ao / FPSO}
Engenheiro de Produ\c{c}\~ao, Process Field Engineer, Facilities Engineer, Operations Engineer, FPSO Cargo Supervisor.

\subsection{Perfura\c{c}\~ao}
Drilling Engineer, Directional Driller Specialist, MPD Supervisor, Project Drilling Engineer, Graduate Trainee.

\subsection{Canais}
LinkedIn (configurar alertas), Jobartis, Careerjet Angola, AngoVagas, p\'aginas de carreira das empresas e recrutadoras (Airswift, CA Global).

\section{Prepar\'acao para candidaturas e entrevistas}
\begin{itemize}
  \item Organizar pasta digital com CV (PT/EN), certificados, diploma e refer\^encias.
  \item Aplicar em ondas: 10--15 candidaturas por ciclo; acompanhar em 7--10 dias.
  \item Preparar exemplos t\'ecnicos (mini-projeto no GitHub) para demonstra\c{c}\~ao pr\'atica.
  \item Treinar respostas comportamentais com STAR (Situa\c{c}\~ao, Tarefa, A\c{c}\~ao, Resultado).
\end{itemize}

\section{Escada de carreira detalhada}
\subsection{Produ\c{c}\~ao}
\begin{itemize}
  \item Trainee / Estagi\'ario: 0--12 meses -- Medical, BOSIET, H2S, IWCF L1 (usd equivalentes variam conforme empresa).
  \item J\'unior: 1--3 anos -- OPITO Processing (aumenta empregabilidade).
  \item Pleno: 3--6 anos -- NEBOSH IGC recomendado.
  \item S\'enior / L\'ider: 6--10 anos -- especializa\c{c}\~oes FPSO.
\end{itemize}

\subsection{Perfura\c{c}\~ao}
\begin{itemize}
  \item Trainee: 0--12 meses -- base (Medical, BOSIET, H2S, IWCF L1).
  \item J\'unior/Field Engineer: 1--3 anos -- IWCF Level 2.
  \item Pleno/S\'enior: 3--10 anos -- IWCF L3/L4, MPD; progress\~ao r\'apida com experi\^encia em campanha.
\end{itemize}

\section{Mini-projetos pr\'aticos (portf\'olio) -- prioridades e ferramentas}
Sugest\~oes de projetos com alto impacto para entrevistas:
\begin{enumerate}
  \item Decline Curve Analysis App (Streamlit + Plotly): upload de CSV, ajuste de curvas, cen\'arios de previs\~ao.
  \item Dashboard de Performance de Po\c{c}os (Power BI): KPIs, anomalias, drill-down por po\c{c}o.
  \item Nodal Analysis Interativo (Python): calcula ponto de opera\c{c}\~ao e sensitivity analysis.
  \item Predi\c{c}\~ao de ROP (XGBoost): modelo ML com feature engineering sobre dados abertos.
  \item Checklist digital de inspe\c{c}\~ao (Streamlit): fotos, logs, export CSV/relat\'orio.
\end{enumerate}

Documente tudo no GitHub, inclua README completo, GIFs e um pitch de 2 minutos para cada projeto.

\section{Migra\c{c}\~ao internacional e valida\c{c}\~oes}
\subsection{Pilares de valida\c{c}\~ao}
OPITO BOSIET/HUET, IWCF (levels), NEBOSH IGC, diploma reconhecido e flu\^encia em ingl\^es (TOEIC/IELTS) s\~ao os principais elementos que validam o seu perfil fora de Angola.

\subsection{Roteiro de migra\c{c}\~ao}
\begin{enumerate}
  \item Consolidar experi\^encia local (3--5 anos) em projetos com empresas multinacionais.
  \item Atualizar CV e LinkedIn em ingl\^es; construir network e seguir recrutadoras internacionais.
  \item Procurar vagas com sponsorship e preparar tradu\c{c}\~oes juramentadas do diploma quando necess\'ario.
\end{enumerate}

\section{Pr\'oximos passos inmediatos (checklist)}
\begin{itemize}
  \item Abrir conta de poupan\c{c}a dedicada.
  \item Atualizar CV e LinkedIn (PT/EN); inscrever-se na SPE.
  \item Agendar exame m\'edico para Julho/2026.
  \item Iniciar estudo IWCF Level 1 e criar primeiro mini-projeto (dashboard de produ\c{c}\~ao).
\end{itemize}

\section*{Conclus\~ao}
Este relat\'orio reune todo o trabalho e as decis\~oes praticas: or\'camento, cronograma, lista de certifica\c{c}\~oes, fornecedores, vagas, mini-projetos e plano de migra\c{c}\~ao internacional. Execute os passos em ordem, mantenha disciplina de poupan\c{c}a e prove continuamente suas habilidades com projetos pr\'aticos. Boa sorte — avise-me quando quiser que eu gere o PDF final e fa\c{c}a o commit e push das altera\c{c}\~oes.

% ------------------------- Fim do Relat\'orio Inserido -------------------------

\end{document}
