\documentclass[openany,11pt]{book}
\input{chain-style.tex}

% ══════════════════════════════════════════════════════════════════
%  LOAD REGIONAL LANDSCAPE SYSTEM  (280 variations)
% ══════════════════════════════════════════════════════════════════
\input{landscape-regions-complete.tex}

% ══════════════════════════════════════════════════════════════════
%  LOAD REGIONAL CHARACTER GLYPH SYSTEM  (180+ variations)
% ══════════════════════════════════════════════════════════════════
\input{character-glyphs-complete.tex}

% ══════════════════════════════════════════════════════════════════
%  REGIONAL GLYPH RESOLVER
%  Wrapper that uses CharacterRegionEmotion context to display correct glyph
% ══════════════════════════════════════════════════════════════════
\newcommand{\glyphRegional}{\POVGlyphRegional}

\title{Embers of the Forsaken Keep}
\author{Your Name}
\date{\today}

\begin{document}

% Configure the book
\SetTotalChapters{3}
\setTopLocation{The Grey Mountains · Dusk}
\setBottomLocation{Valley of Ashenveil · Twilight}

% ═══════════════════════════════════════════════════════════════════════════
% CHAPTER I
% ═══════════════════════════════════════════════════════════════════════════

\POVchapter{A}{Part I: The Grey Mountains}

\setPOVName{AELINDRA}

\lettrine[lines=3]{T}{he wind came down from the Grey Mountains carrying the smell of burnt cedar and something older—something that had no name in the common tongue, but which every soldier recognised in the hollow of their ribs.} Aelindra tightened the strap of her vambrace and said nothing.

Below, the valley curved like a wound in the earth, its floor dark with the remnants of what had been three villages not a fortnight past. The smoke still rose in thin, patient threads, as though the ground itself had not yet decided whether to forgive. She counted the fires. Seven. No—eight, if the far glow past the treeline was theirs and not some shepherd's accident.

Eight fires meant a column, not a raid. A column meant orders. Orders meant someone, somewhere, had decided this particular valley was worth the cost of making it into a memory.

\scenebreak

Aelindra had spent seventeen years learning the difference between decisions made in rooms and decisions that lived in the soil afterward. She had yet to find a word for the distance between them.

``We should not have come this way,'' said the boy behind her. She did not turn. The boy said many true things. That was what made him dangerous to keep near and impossible to send away.

``The path chooses itself,'' she said. It was not an answer, but the boy was learning to accept answers that were not answers at all.

\begin{quotebox}
  \textit{``The path chooses itself. The traveller merely walks where choice has led them. Such is the burden of command—to walk the path and pretend it was never chosen at all.''}
  
  \hfill—Aelindra's Journal, Third Year of the War
\end{quotebox}

When the boy began speaking again, she simply raised her hand. Some silences were deeper than words. The smoke rose. The wind carried ash and the scent of endings. And Aelindra, who had given all her years to preventing catastrophe, watched it unfold anyway in the valleys below.

\newpage

% ═══════════════════════════════════════════════════════════════════════════
% CHAPTER II
% ═══════════════════════════════════════════════════════════════════════════

\POVchapter{R}{The Scout's Return}

\setPOVName{RAVEN}

\lettrine[lines=3]{T}{he courier horses were near exhaustion when they came pounding down the mountain pass.} Raven watched from the outcropping, counting them: three riders, all in the colours of the western garrison, all riding hard enough to kill their mounts. That meant news. Bad news. The kind that made commanders ride for three days straight without rest.

He slipped down the rock face with the ease of long practice, moving through shadows like water. The horses hadn't seen him. The riders wouldn't either until he wanted them to.

The scout's trade was learned slowly and forgotten quickly once you stopped paying attention. Raven wasn't the sort to stop paying attention.

\scenebreak

By sunfall, he would have their message. By the following dawn, he would know everything they knew, and perhaps a few things they didn't. That was why Aelindra kept him close, despite the other captains' misgivings. A good scout could see around corners. A great scout could see around time itself.

Raven was, by his own estimation, somewhere in between.

The lead rider's horse stumbled. Raven found himself frowning at that—they might have to walk the last three miles if the animal went lame. That would delay the message. That might change everything.

\begin{quotebox}
  \textit{``In the end, it's never the sword that matters. It's what you know, and when you know it, and whether you're willing to act on knowledge come what may.''}
  
  \hfill—A saying among the scouts
\end{quotebox}

He stayed where he was, hidden, watching. Already his mind was weaving possibilities: if the news was what he suspected, then Aelindra would need to move the garrison by nightfall. If she moved the garrison, the supply lines would break for a day at least. If the supply lines broke...

The chain of consequence was long. Raven had learned to follow it far enough to matter.

\newpage

% ═══════════════════════════════════════════════════════════════════════════
% CHAPTER III
% ═══════════════════════════════════════════════════════════════════════════

\POVchapter{☽}{The Mystic's Vision}

\setPOVName{SEREN}

\lettrine[lines=3]{S}{eren saw the world in layers—the physical world of stone and soil and burning wood, and beneath that, the world of story, the world of meaning, the world that most people stumbled through blind.} In the tower room, sitting before the scrying pool, she saw both layers at once.

What came was not a vision like the ones the old texts described. It was subtler. It was the way the light shifted on the water's surface. The way a single ripple seemed to contain a whole geometry of consequences.

Something had changed. The weight of the world had shifted, almost imperceptibly, but she felt it in the root of her skull, in the deepest parts of her gift.

The garrison would move. That much was clear in the way the future seemed to thin and fold. Aelindra would move them. And in moving, she would unweave something that had been woven very carefully, very deliberately, over nearly a generation.

\scenebreak

Seren rose from the pool and steadied herself against the stone rim. The visions never left her whole. They took pieces—a moment of clarity for a week of headaches, a glimpse of tomorrow for a day of muddled thinking.

She walked to the tower window and looked out over the valley. From here, in the thin place where the veil wore worn, she could see the smoke from eight fires.

But she could also see the shadows moving beneath them. The cold gathering at the edges of things. The breath of something vast and patient, waiting in the spaces between.

\begin{quotebox}
  \textit{``The gift of sight is also the curse of knowing. Once you have seen what comes, you cannot unsee it, and you cannot prevent it. You can only choose how you stand when it arrives.''}
  
  \hfill—Seren's Testament
\end{quotebox}

The boy would come to the tower by nightfall. Everything hinged on what the boy would choose to do when he understood the full truth of what had been set in motion. For now, she could only wait, and watch, and carry her knowledge like a stone in her chest.

\end{document}
