\documentclass{cwbook}

\begin{document}

\BookTitle{Ashes of the Weave}{A sample layout for POV-driven dark fantasy}

\tableofcontents
\clearpage

\chapter{The Broken Accord}
\ENV{city}

\POV{\GlyphAisen}{Aisen}

Aisen stood beneath the arches of Aldric's Crossing while the rain turned the lanternlight to bruised gold. He had learned, by bitter repetition, that every bargain carried a blade hidden inside it.

Caspian Vane waited at the far end of the bridge, hands folded, his expression almost gentle.

\ShiftCue[The thread passes by touch.] 

\POV{\GlyphCaspian}{Caspian Vane}

Caspian watched Aisen approach and measured the silence between them like a wound that could still be reopened. He had already decided that this meeting would cost him something.

Aisen spoke first. The words were careful, but the kind of careful that hid steel.

\SceneBreak

They crossed the bridge together and found Ser Corvin Ashford in the chapel ruin beyond the market ward.

\ShiftCue[The witness changes.] 

\POV{\GlyphCorvin}{Ser Corvin Ashford}

Corvin had been waiting long enough for grief to become architecture. When he looked at them, it was not surprise that shaped his face, but recognition of a future already in motion.

He turned toward the stairwell leading into the sealed undercroft, and the story went with him.

\chapter{Under the Stone}
\ENV{dungeon}

\POV{\GlyphAisen}{Aisen}

The undercroft swallowed sound. Aisen followed the torchlight down into the old vaults, each step carrying him deeper into the places where the academy had buried its oldest mistakes.

Ryo appeared from the dark as if lightning had decided to learn a human shape.

\ShiftCue[Consequences first, explanations later.] 

\POV{\GlyphRyo}{Ryo}

Ryo had already crossed the chamber by the time Aisen understood that he was not a rescue but a fault line. The air around him crackled with a force that did not belong to this world alone.

He looked past Aisen, toward the broken gate.

Kael was there. Or had been, only moments before.

\ShiftCue[We follow the one who carries the next turn.] 

\POV{\GlyphKael}{Kael}

Kael had entered the vault to save someone and discovered, too late, that salvation was a narrower corridor than punishment. Behind him, the stone shuddered.

Ahead, something old began to wake.

\chapter{The Loop Beneath Ice}
\ENV{northernwastes}

\POV{\GlyphKairos}{Kairos}

Kairos moved through the frozen pass as though time itself were something he had already broken and was now carrying in pieces. He had seen the same night return too many times to trust it, yet still he walked forward.

Orin emerged from the snow like a memory that refused to stay buried.

\ShiftCue[The abandoned perspective becomes visible.] 

\POV{\GlyphOrin}{Orin}

Orin did not understand why the wind sounded familiar. He only knew that the man before him looked like he had died in a hundred different ways and returned with none of them fully gone.

Far away, the Covenant fires moved over the ridge.

\chapter{Ash in the Wastes}
\ENV{wasteland}

\POV{\GlyphAmara}{Amara}

Amara had spent years building a legacy strong enough to survive betrayal, and now she stood at the edge of a dead plain watching it come apart piece by piece.

Renard waited beside her, quiet and ancient, as if he had already outlived the argument.

\ShiftCue[When the hidden witness speaks, the shape of the truth changes.] 

\POV{\GlyphRenard}{Renard}

Renard had seen empires fail for less than this. He looked at the horizon and knew that the next move would not be made by armies, but by the frightened choices of three people who had not yet learned how much power could hide inside silence.

\end{document}
